---
title: 'Contribution of RPAS in research and conservation in protected areas: present
  and future'
author:
- affiliation: University
  name: Jesús Jiménez López
- affiliation: University
  name: Margarita Mulero-Pázmány
bibliography:
- ./internet.bib
- ./master.bib
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    pandoc_args:
    - -V
    - classoption=onecolumn
    template: ./svm.latex.ms.tex
  html_document: default
csl: ./taylor-and-francis-chicago-manual-of-style-author-date_with_note.csl
documentclass: article
keywords: RPAS, UAV, drones, natural protected areas, conservation, biodiversity
numbersections: yes
cls: ./interact.cls
runhead: null
thanks: null
abstract: In the last few decades, we have witnessed a growing interest in projects
  aimed to evaluate the feasibility of RPAS for conservation purposes. So far, RPAS
  have been proposed  for a variety of research and management activities including
  environmental and wildlife monitoring or law enforcement. It remains to be seen
  how well RPAS complement conservation actions in protected areas, since beyond technical,
  ethical and legal barriers that compromise their effectiveness,  it is a common
  source of discussion that bias in research efforts is often of limited utility in
  addressing practical issues, particularly those claimed by natural park managers.
  We carried out a systematic review of the published literature in order to consider
  whether more attention should be given to certain conservation activities where
  the contribution of RPAS could improve the efficiency in the management of protected
  areas. After discussing results, we expose main drawbacks and identify those areas
  that requires further research investment, highlighting some trends and opportunities
  that apparently have not yet been adequately exploited.
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

As defined by UICN, "a protected area is a clearly defined geographical space, recognized, dedicated and managed, through legal or other effective means, to achieve the long term conservation of nature with associated ecosystem services and cultural values" [@dudley_guidelines_2008]. Protected areas have been declared under different reasons and circumstances but there is a general consensus on its importance in safeguarding biodiversity, contribute to  human well-being and ensure persistece of  the natural heritage for the enjoyment of future generations.  Despite such praiseworthy intentions, the reality faced by protected areas is subject to a wide variety of requirements and unforeseen challenges requiring rapid and effective solutions.

To ensure its ability to meet both long-term and short-term conservation goals, protected areas have benefited from a extensive regulatory framework, to a certain extent implemented through the management plans. These plans aimed to establish the basic guidelines to protect the natural and socioeconomic values by which protected areas were declared. In the field of competences derived from such instruments, available human and material resources have mainly been allocated to regulate tourism and recreational activities, monitoring law enforcement including various forms of illegal resource extraction, wildfire prevention and fighting, wildlife monitoring campaigns to maintain up-to-date fauna inventories, environmental assesment or actions aimed at strengthen environmental education and research programs. Those activities have benefit from a wide range of technological advances, including remote sensors, field monitoring stations, manned surveys, camera traps, wildlife tracking devices or computational tools. More recently, applications of remotely piloted aircraft  systems (RPAS, also known as unmanned aerial systems, UAS, drones) have been the subject of a growing interest in both the civilian sphere and scientific community. Not surprisingly, there have been a significant amount of  research projects facing conservation issues using RPAS, especially during the last decade [@linchant_are_2015; @chabot_wildlife_2015; @christie_unmanned_2016]. As a consequence, RPAS have been suggested as an appropriate complement for conservation activities [@koh_dawn_2012; @rodriguez_eye_2012; @chabot_wildlife_2015, @wulder_high_2004, @zahawi_using_2015].

<!-- ojo repetido chabot_wildlife_2015 -->

However, to date, it has not been adequately considered whether research efforts using RPAS meet the demands of conservation practitioners.  The fact that the RPAS has received increasing attention in studies addressing its feasibility to approach ecological issues and practical conservation problems does not necessarily respond to the priorities raised. In addition, budgetary constraints are common in protected area management and available resources are generally insufficient for optimal compliance. While park managers are responsible for acquiring as much information as possible to ensure that such potential source of investment will have a positive effect on the management efficiency and quality of services provided, the commitment of the research community is to provide the appropriate responses to facilitate better decision-making. We note that wildlife and habitat monitoring has received considerable attention. However, the scarcity of lines of research in other areas of equal importance is an important gap to consider when attempting to demonstrate the true potential of RPAS, encouraging its formal application in protected area management tasks.

However, to date, it has not been adequately considered whether research efforts using RPAS meet the demands of conservation practitioners.  As noted, wildlife and habitat monitoring have received major emphasis. Consequently, this effort does not necessarily respond to practical conservation priorities. In addition, budgetary constraints are common in protected area management and available resources are generally insufficient for optimal compliance. Park managers must have sufficient guarantees to ensure that such potential source of investment will have a positive effect on the management efficiency and quality of services provided. The commitment of the research community is to provide the appropriate responses to facilitate better decision-making. Fostering RPAS in protected areas requires cover this important gap to consider when attempting to demonstrate the true potential of RPAS, encouraging its formal application in protected area management tasks.

, while there are important gaps to consider when uncovering the true potential of RPAS for conservation tasks.

There is an uneven amount of studies facing 

# Methods

To achieve the proposed objectives, a systematic bibliographical review (see PRISMA Flowchart) of scientific articles, gray literature, postgraduate theses, websites and specialized journals was carried out, following a similar line to other related studies [@linchant_are_2015; @christie_unmanned_2016; @mulero-pazmany_unmanned_2017].  Last reference revised was published on June, 2017. The main tools for selecting bibliography include Google Schoolar, Research Gate and Mendeley Desktop, while the use of Internet search engines include other references outside the scientific scope. Key search criteria, primarily in English, include RPAS in their various meanings and acronyms, reflecting the varied terminology used: "remotely piloted aerial system (RPAS)", "drone", "unmaned aircraft system (UAS)" "unmaned aircraft", "model aircraft", "radio control aircraft" , "unmaned aerial vehicle (UAV)" along with several terms referring to common activities carried out in protected areas, from the general to the particular: "conservation", "protected areas", "ecology" "ecosystem", "habitat", "vegetation", "forest", "wetland", "reforestation", "monitoring", "survey",  "sampling", "inventory", "wildlife", "fauna", "bird", "mammal", "fish, "amphibian", "reptile", "wildfire", "remote sensing", "tourism", "ecotourism", "law enforcement", "poaching", "logging", "risk management", "pollution" and so on. For example, a restricted search was carried out with a logical statement as follows: "conservation drone" OR "conservation UAS" OR "conservation UAV" OR "conservation unmanned aircraft system" OR "conservation unmanned aircraft" OR "conservation RPAS" OR "conservation radio control aircraft", both in singular and plural forms, as there are limits in the number of words in a single search statement. That counts for more than 80 posible combinations. We also included references from above mentioned review articles and those not indexed in research results but relevant for the purpose of this study.

From the total number of publications (x), we excluded duplicated results or those with no relevance for conservation, based on activities included in protected area management plans worldwide. The selected publications (x) were categorized according to the main purpose of the study. These categories include: "wildlife monitoring and management", for those feasibility studies facing alternative population surveys and fauna tracking methods, together with advances in software development focused on automatic animal counting and detection; "monitoring and mapping of terrestrial and aquatic ecosystems", for vegetation surveys, ecosystem dynamics studies and habitat characterization; "Law enforcement" encompasses poaching, illegal logging and fishing activities; "ecotourism" refers mainly to recreational activities and visitors management; "risk management, enviromental assesment and decision support" encompass a wide variaty of duties, including pollution sampling, disaster management. Finally, legal constraints and actions to minimize impact on fauna are also considered, as both shape the feasibility of RPAS to approach conservation and environmental issues.

The collected information  is presented in tabular format, identifying where the study was conducted, the expected accomplishments and technical specifications of the aerial platform. After expose main results, gaps are identified and possible scenarios for implementing  RPAS as essential tools to help achieve conservation goals in protected areas are discussed.

<!-- include cost or potential sources of costs -->

# Results

## Wildlife Monitoring and Management

Wildlife science generally refers to the study, monitoring, and management of wild mammals, birds, reptiles, and amphibians (and occasionally fish), as well as their habitats.

Several studies focused on develop survey methods aimed at locating, counting and characterizing wildlife communities and habitats, ranging from  large and medium size terrestrial mammals [@lancia_estimating_2005; @jain_unmanned_2013, @barasona_unmanned_2014],  birds [@wilson_feasibility_2017, @hodgson_precision_2016,@christie_unmanned_2016 @sarda-palomera_fine-scale_2012, @chabot_evaluation_2012], but also species relying on coastal and marine ecosystems  [@colefax_potential_2017; @hodgson_unmanned_2017; @_testing_2016; @koski_evaluation_2015, @dulava_environmental_2015, @durban_photogrammetry_2015, @koski_evaluation_2009]. RPAS have as well been applied as a means of inspecting and assess  breeding and nesting areas at inaccessible sites in several species [@szantoi_mapping_2017; @wich_preliminary_2016, @puttock_aerial_2015; @van_andel_locating_2015; @weissensteiner_low-budget_2015 ] or as a complement for wildlife telemetry tracking methods [@christie_unmanned_2016; @bayram_active_2016; @mulero-pazmany_unmanned_2015; @korner_autonomous_2010; @cliff_online_2015; @leonardo_miniature_2013; @soriano_rf-based_2009;]. Given the large amount of information generated, it is not surprising that computer vision software have been developed to handle the automatic detection, recognition and counting of individuals captured in scenes acquired by visible and thermal-infrared sensors, replacing otherwise time-consuming manual tasks  [@andrew_semi-automated_2017; @chabot_computer-automated_2016; @gonzalez_unmanned_2016, @lhoest_how_2015; @van_gemert_nature_2015;  @christiansen_automated_2014; @martin_estimating_2012; @abd-elrahman_development_2005]. 


\newpage
\blandscape

\tiny
```{r  echo=FALSE}

df <- read.csv("tabla.csv", header=FALSE)
library(kableExtra)
library(pander)
library(knitr)
# x <- kable(df, format = "pandoc", caption = "Title of the table",  booktabs = TRUE)
# column_spec(x, 1, width = "1cm", bold = TRUE, italic = TRUE)

```

| Study | Aims | Country | Place | Target | RPAS platform | Payload | Costs |
|------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------------------|------------------------|--------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------|----------|
| [@mulero-pazmany_unmanned_2015] | Telemetry/RPAS SDM comparative study | Spain |  Doñana National Park | *Bos taurus* | Fixed-wing: Easy Fly plane, Ikarus autopilot, Eagletree GPS logger | Panasonic Lumix LX-3 11MP | $ 6500 |
| [@hodgson_unmanned_2017] | Comparative survey RPAS/land based observation; abundance estimation | Australia | North Stradbroke Island | humback whales | Fixed-wing: ScanEagle | Nikon D90 12MP, Standard Definition Electro-Optical Camera | ? |
| [@hodgson_unmanned_2013] | Dugongs detection | Australia | Shark Bay Marine Park | Dugong | Fixed-wing: ScanEagle | Nikon D90 12MP | ? |
| [@wilson_feasibility_2017] | Bioacustic monitoring | USA | State Game Lands | Birds | Rotor-wing: DJI Phantom 2 | ZOOM H1 Handy Recorder | ? |
| [@bayram_active_2016] | VHF collars tracking | ? | ? | Bears (Ursus) | Rotor-wing: DJI F550 | Telonics MOD-500 VHF, Uniden handheld scanner | ? |
| [@christie_unmanned_2016] | Abundance estimation | USA | Aleutian Islands | Steller Sea Lion (Eumetopias jubatus) | Rotor-wing: APH-22 | ? | $ 25.000 |
| [@christie_unmanned_2016}] | Abundace estimation | USA | Monte Vista National Wildlife Refuge | Grus canadensis (sandhill cranes) | Fixed-wing: Raven RQ- 11A | ? | $ 400 |
| [@wich_preliminary_2016] | Sumatran orangutan nest detection | ? | ? | ? | Fixed-wing: Skywalker 2013 | Canon S100 | ? |
| [@van_andel_locating_2015] | Chimpanzee nest detection | Africa | Loango National Park | Chimpanzee (Pan troglodytes) | Fixed-wing: Maja | Canon Powershot SX230 HS | $ 5000 |
| [@koski_evaluation_2009] | Marine mammals monitoring | USA | Admiralty Bay | Marine mammals | Fixed-wing: ScanEagle | NTSC Video Camera | ? |
| [@andrew_semi-automated_2017] | Semi-automated image processing tools to detect and map sea eagle nests | Australia | Houtman Abrolhos Islands | White-bellied sea eagle (Haliaeetus leucogaster) | ? | ? | ? |
| [@longmore_adapting_2017] | Software development to help detect animals in thermal images | UK | Arrowe Brook Farm Wirral | Wildlife | Rotor-wing: 3DR robotics Y6 | FLIR, Tau 2 LWIR Thermal Imaging Camera Core | ? |
| [@martin_estimating_2012] | Estimate the distribution of organisms using statistical models | USA | ? | Manatee (*Trichechus manatus latirostris*) | Fixed-wing: Nova 2.1 | Olympus H E-420 | - |
| [[@chabot_evaluation_2012]] | Waterbirds surveys | Canada | Ste-Anne-de-Bellevue, Baie-du-Feb- vre | Canada Geese (*Branta canadensis*), Snow Geese (*Chen caerule- scens*) | Fixed-wing; CropCam | Pentax Optio A20 10MP | $ 11000 |
| [@koski_evaluation_2015] | Reidentification of photographed whales  | Canada | Igloolik Island | Bowhead whales (*Balaena mysticetus*) | Fixed-wing: TD100E | GoPro, Nikon D800 50 mm f1.2 Nikon lens | - |
| [@durban_photogrammetry_2015] | Photogrammetry studies  | Canada | Vancouver Island | Killer whales (*Orcinus orca*) | Rotor-wing: APH-22 | Olympus E-PM2 Olympus M.Zuiko 25 mm F1.8 lens | - |
| [@lhoest_how_2015] | Develop algorithm for automatic count of hippos  | Congo | Garamba National Park | Common hippopotamus (*Hippopotamus amphibius L.*) | Fixed-wing: Falcon | Tamarisk 640 | - |
| [@goebel_small_2015] | Estimating abundance and size of Antarctic predators  | Antarctica | Cape Shirreff, Livingston Island, South Shetland Islands | Antarctic fur seal (*Arctocephalus gazella*), Penguin (*Spheniscidae*), Leopard seal (*Hydrurga leptonyx*) | Rotor-wing: md4-1000, APQ-18, APH-22 | Sony NEX-5, Canon EOS-M,Olympus E-P1 | - |
| [@dulava_environmental_2015] | Waterbirds surveys | USA | Ruby Lake NWR, Tomales Bay |mallard (*Anas platyrhynchos*), northern shoveler (*Anas clypeata*), gadwall (*Anas strepera*), blue-winged teal (*Anas discors*), northern pintail (*Anas acuta*), American wigeon (*Anas americana*), canvasback (*Aythya valisineria*), scaup (*Aythya marila/ Aythya affinis*), surf scoter (*Melanitta perspicillata*), American coot (*Fulica americana*) | Rotor-wing: Honeywell RQ-16 T-Hawk , Fixed-wing: AeroVironment RQ-11A | Canon PowerShot SX230, SX260 12.1MP, GoPro Hero3 | - |
[@cliff_online_2015] | Online Localization of Radio-Tagged Wildlife | Australia | - | Noisy miner (M*anorina Melanocephala*) | Rotor-wing: Falcon 8 | Radiometrix LMR1 receiver, Digi XTend radio modem | - |
[@_testing_2016] | Salmon spawning surveys | USA | Lower Snake River | Chinook salmon (*Oncorhynchus tshawytscha*) | Rotor-wing: Aeryon Scout | GoPro Hero3 | - |
[@barasona_unmanned_2014] | Epidemiological study  | Spain | Doñana National Park |  Red deer (*Cervus elaphus*), Fallow deer (*Dama dama*), Cattle (*Bos taurus*) | Fixed-wing: Easy Fly plane | Panasonic Lumix LX-3 11MP | - |
[@jain_unmanned_2013] | Elephant surveys  | Burkina Faso | Nazinga Game Ranch | Elephant (*Loxodonta africana*) | Fixed-wing: Gatewing x100 | Ricoh GR III  | - |
[@rodriguez_eye_2012] | UAS and GPS Data Loggers | Spain | Doñana National Park  | Lesser kestrel (*Falco naumanni*) | Fixed-wing: ST-model Easy Fly | Panasonic Lumix LX-3 | - |
[@sarda-palomera_fine-scale_2012] | Fine-scale bird monitoring | Spain | Estany d’Ivars i Vila-Sana | Black-headed (*Gull Chroicocephalus ridibun*) | fixed-wing: Multiplex Twin Star II | Panasonic Lumix FT-1 12MP 28–128 mm  | $ 1645 |
[@hodgson_precision_2016] | Ccolony nesting birds surveys | Ashmore Reef Commonwealth Marine Reserve, Adele Island, Macquarie Island | Australia | Frigatebirds (*Fregata ariel*), Crested Terns (*Talasseus bergii*), Royal Penguins (*Eudyptes schlegeli*)  | Rotor-wing: X8 3D Robotics, Fixed-wing: FX79 | Canon EOS M 40mm | $ 5100 |

[@-] | - | - | - | - | - |  | - |


Table: WILDLIFE MONITORING AND MANAGEMENT

| Study | Aims | Country | Place | Target | RPAS platform | Payload | Costs |
|------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------------------|------------------------|--------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------|----------|
| [@perroy_assessing_2017] | Tropical invasive plants | USA | Pahoa, Hawai | *Miconia calvescens* | Rotor-wing: DJ Inspire-1 | DJI FC350 camera | ? |
| [@szantoi_mapping_2017] | Habitat Mapping | Indonesia | Gunung Leuser National Park | Orangutan (*Pongo abelii*) | Fixed-wing: Skywalker | Canon S100 | $ 4000 |
| [@casella_mapping_2017] | Coral reef mapping | French Polynesia | Tiahura,; Moorea | Coral reef | Rotor-wing: DJI Phantom 2 | Modified GoPro HERO4 | $ 1678 |
| [@casella_drones_2016] | Monitoring coastal erosion dynamics in shorelines | French Polynesia | Tiahura; Moorea | Coral reef | Rotor-wing: Mikrokopter Okto XL | Canon G11 | $ 7500 |
| [@mullerova_does_2016] | Monitoring plant invasion | ? | ? | Exotic species | Fixed-wing: VUT 712 713  720 | Canon S100 | ? |
| [@ventura_low-cost_2016] | Marine fish nursery areas mapping | Italy | Giglio Island | Marine fish nursery areas | Rotor-wing: homemade prototype | Mobius HD, GoPro HERO3 Black Edition | $ 100 |
| [@ivosevic_use_2015] | Habitat monitoring and modeling in restricted areas; RPAS performance test & South Korea & Chiaksan National Park;Taeanhaean National Park | South Korea | Chiaksan National Park;Taeanhaean National Park |  | Rotor-wing: DJI Phantom 2 Vision+ | Full HD videos 1080p/30fps and 720p/60fps | ? |
| [@lisein_discrimination_2015] | Discrimination of deciduous species; Forest inventory | Belgium | Grand-Leez | English oak, birches, sycamore maple ,common ash and poplars | Fixed-wings: Gatewing X100 | Ricoh GR2 GR3 GR4 10 megapixels CCD | ? |
| [@puttock_aerial_2015] | Characterization of ecosystems affected by beaver activity | UK | Devon Beaver Project site | Eurasian beaver (*Castor fiber*) | Rotor-wing: 3D Robotics Y6 | anon ELPH 520 HS | ? |
| [@zahawi_using_2015] | Characterization of tropical forest structure for restoration actions | Costa Rica | Devon Beaver Project site | Several species | Rotor-wing: 3D Robotics Y6 | Canon S100 | $ 1500 |
| [@bustamante_forest_2015] | Forest monitoring | Brasil | Riverine Forests (Permanent Protected Areas), Rio de Janeiro, Barrãcao do Mendes, Santa Cruz and São Lorenço | Riverbank forests | Rotor-wing: DJI Phantom Vision 2S | RGB digital camera 14MP | $ 9700 |
| [@gini_aerial_2012] | 3D modeling and classification of tree species | Italy | Parco Adda Nord | Several species | Rotor-wing: Microdrones TM MD4-200 | RGB CCD 12 megapixels Pentax Optio A40; modified NIR Sigma DP1 with a Foveon X3 sensor | ? |
| [@miyamoto_use_2004] | Classification of species in wetlands | Japan | Kushiro Wetlands | Several species | Helium balloon | NIKON F-801, NIKKOR 28 mm f/2.8 | $ 1600 |
| [@casella_mapping_2017] | Mapping coral reefs | ? | ? | ? | Rotor.wing: DJI Phantom 2 | GoPro HERO4 | ? |

Table: MONITORING OF TERRESTRIAL AND AQUATIC ECOSYSTEMS 


| Study | Aims | Country | Place | Target | RPAS platform | Payload | Costs |
|------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------------------|------------------------|--------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------|----------|
| [@lobermeier_mitigating_2015] | Mitigate the risk of collision by installing markers on electrical lines | USA | ? | Birds | Rotor-wing: Mikrokopter Hexa XL | KX 171 Microcam | ? |
| [@margarita_mulero-pazmany_juan_jose_negro_low_2014}] | Bird risk hazards in power lines | Spain | Doñana National Park | Birds | ixed-wing: Easy fly St-330 | GoPro HERO2 11MP;Panasonic LX3 11MP | $ 8863 |
| [@mulero-pazmany_remotely_2014] | Anti-poaching | Africa | KwaZulu-Nata | Black rhinocero (*Diceros bicornis*), white rhinocero (*Ceratotherium simum*) | Fixed-wing: Easy Fly St-330 | Panasonic Lumix LX-3 11 MP, GoPro Hero2; Thermoteknix Micro CAM microbolometer | $ 15700 |
| [@hansen_applying_2016] | Visitors Surveillance | Sweden | Sweden & Kosterhavet National Park | Humans | ? | ? | ? |
| [@sabella_preliminary_2017] | Visitors Surveillance | Italy | R.N.O. Oasi faunistica di Vendicari | Humans | Rotor-wing: DJI Phantom 3 | ? | ? |
| [@king_will_2014] | RPAS applications in ecotourism activities | Sweeden | Sweeden & Kosterhavet National Park | Humans | ? | ? | ? |
| [@vas_approaching_2015] | RPAS impact | France | Zoo du Lunaret, Cros Martin Natural Area | *Anas platyrhyncho*, *Phoenicopterus roseus*, *Tringa nebularia* | Rotor-wing: Phantom | GoPro HERO3 | ? |
| [@weissensteiner_low-budget_2015] | RPAS Impact | Sweeden | ? | Hooded crow (*Corvus corone cornix*) | Rotor-wing: DJI Phantom 2 Vision | ? | $ 1000 |

Table: INFRASTRUCTURES AND RISK ASSESSMENT, ECOTOURISM, IMPACT ON WILDLIFE AND ECOSYSTEMS 

| Study | Aims | Country | Place | Target | RPAS platform | Payload | Costs |
|------------------------------------------------------------------------------------|--------------------------------------------------------------------------------------------------------------------------------------------|------------------------|--------------------------------------------------------------------------------------------------------------|-------------------------------------------------------------------------------|--------------------------------------------------------------------|----------------------------------------------------------------------------------------|----------|
| [@zang_investigating_2012] | Pollution monitoring | China | Several cities | Rivers | Fixed-wing | Canon 50D, ACD multispectral camera | ? |
| [@cornell_use_2016] | Water sampling | USA | Lake Ontari0 | ? | Rotor-wing: DJI Phantom 3 | 50mL Falcon tube | ? |
| [@mccaldin_use_2015] | Aerial baiting | Australia | Christmas Island | Cat (*Felis catus*) | Rotor-wing: V-TOL Hornet I-II | Canon S100; Drop mechanism with HD Video Recorder | ? |
| [@fornace_mapping_2014] | Spatial epidemiology | Malaysia / Philippines | Sabah / Palawan | ? | Fixed-wing: Sensefly eBee | 16mp | $ 25000 |
| [@van_tilburg_first_2017] | Search and Rescue (SAR) | USA | Columbia Gorge National Scenic Area | Humans | Rotor-wing: Phantom 3, SAR Bot,  Inspire 1 | DJI 12MP; VUE PRO 640 thermal imager | ? |
| [@schwarzbach_remote_2014] | Water sampling | Spain |  Doñana National Park | Freshwater ecosystems | Rotor-wing: Helicopter | Water sampling mechanism | ? |
| [@schmale_development_2008] | Aerobiological sampling | USA | Virginia Tech’s Kentland Farm | Prokaryotic and eukaryotic microorganisms | Fixed-wing: Senior Telemaster | Aerobiological sampling devices | ? |


Table: ENVIRONMENTAL MONITORING AND DECISION SUPPORT

<!-- Table: RPAS APPLICATIONS IN  NATURAL PROTECTED AREAS -->

\elandscape
\normalsize

## Monitoring and mapping of terrestrial and aquatic ecosystems

So far RPAS have been used to classify vegetation cover in wetlands  [@USGS2014], monitoring the spread and detection rate of invasive species [@mullerova_does_2016, @zaman_use_2011, @perroy_assessing_2017], the characterization of forest stands [@gini_aerial_2012], including the evaluation of the effectiveness of restoration actions [@zahawi_using_2015] or multitemporal analysis of the spectral response to phenological variations in different species of deciduous trees  [@lisein_discrimination_2015].  Recently, shallow coastal habitats were also mapped using  consumer grade RPAS [@casella_mapping_2017, @ventura_low-cost_2016], but also to monitor erosion dynamics in shorelines. [@casella_drones_2016]. This momentum has been especially notable with the parallel development of affordable multispectral and hyperspectral sensors adapted to small aircraft [@bareth_low-weight_2015]. 

## Infrastructure and risk assessment
  
Other research projects highlight the convenience of RPAS in assessing the risk  that linear electrical infraestructures posed for birds [@margarita_mulero-pazmany_juan_jose_negro_low_2014; @lobermeier_mitigating_2015] or  identify nests on the ground at risk of destruction by harvesting [@mulero-pazmany_margarita_aeromab_2011].

## Law enforcement

RPAS have also  relevance in the control and surveillance of protected areas, documented through different experiences and considerations focused mainly on poaching and illegal fishing activities [@mulero-pazmany_remotely_2014, @franco_five_2016, @olivares-mendez_noah_2014] but also other less contentious illegal activities [@sabella_preliminary_2017]. [@duffy_waging_2014] analyzed the consequences of the militarization of conservation practices as an increasing trend in natural protected areas around the world and illustrates the use of RPAS through several examples.

## Ecotourism

The high degree of diversification offered by RPAS in the ecotourism industry is summarized in a recent article, which shows possible recreational activities, business opportunities, search and rescue operations, mapping and formulas for granting RPAS flight permits  in designated areas  [@king_will_2014]. Within the still scarce literature,  [@hansen_applying_2016] values the effectiveness of RPAS in monitoring visitors in marine and coastal areas, in combination with other innovative solutions. According to the author, RPAS would theoretically allow to operate under different environmental conditions, improving the level of detail and offering a continuous coverage in the flow and behavior of  visitors, as opposed to other techniques of habitual use like the manual observation or the installation of networks of surveillance cameras. 

# Environmental management and decision support

Planning in protected areas is reflected through a variety of management programs that are difficult to fit into some of the previously discussed categories. An extensive report based on Christmas Island proposed dropping poisoned baits from RPAS to eradicate feral cats disturbing threatened native species [@mccaldin_use_2015].  [@cornell_use_2016]  obtained ground truth data by adapting  RPAS to take water samples for comparison with hyperspectral measurements of Landsat 8 Operational Land Imager (OLI).  [@zang_investigating_2012] identified several pollution agents in riparian areas using RPAS imagery. [@schwarzbach_remote_2014] goes further away by performing several aerial water sampling methods using an unmaned helicopter to monitor water pollution, while [@schmale_development_2008] collected a broad spectrum of both   prokaryotic and eukaryotic microorganisms using a fixed-wing aircraft equiped with a custom made aerial sampling device.  [@fornace_mapping_2014] considered mapping enviromental risk factors for predicting zoonotic diseases as part of a extensive epidemiological study carried out in Philippines and so on. Literature citing RPAS for search and rescue activities  is profuse and an in-depth revision is beyond the scope of this article, but a recent publication ilustrate several examples where RPAS were succesfully operated to assist rescue teams [@van_tilburg_first_2017]. A google scholar search sorted by relevance using disaster management and drones keywords throws at first place a complete report describing a complex framework for decision support using RPAS [@maza_experimental_2011]. 

## Impact of RPAS on wildlife and ecosystems

Animal welfare in wildlife management  practices and ecological research is a sensitive issue from which ethical issues arise [@wilson_measuring_2006]. Not surprisingly, RPAS are not exempt of discussion and several trials measure the disturbance effects of RPAS on  birds  [@vas_approaching_2015, @mcevoy_evaluation_2016,@borrelle_will_2017, @scobie_wildlife_2016,  @weissensteiner_low-budget_2015] and mammals [@ditmer_bears_2015, @pomeroy_assessing_2015], while  other studies marginally  inform observed behavioural patterns [@jain_unmanned_2013, @mulero-pazmany_unmanned_2015].  Finally,a code of good practice and recommendations is build upon the learned experiences.

# Discussion 

## Legal barriers
 
RPAS operations faces important legal barriers that undermine the true potential in the civilian sphere  [@stocker_review_2017]. An overly restrictive regulatory framework could limit the possibilities of use of the RPAS in the field of conservation, which makes clear the urgent need to harmonize legislation . In the United States and in most of the European countries consulted, interim legislation has been adopted which, to a certain extent, equates the management of RPAS with that of traditional aircraft. In general terms, the situation in Latin America is uneven, however there is a general tendency to develop specific laws to cope with the rise of the RPAS in both the civil and military sectors [@Bworld_of_drones2017]. Africa is one of the continents where the impact of RPAS in conservation has had greater repercussions. However, in the opinion of some conservationists, their use have not been without problems, resulting in governments that have totally or partially prohibited drone operations, arguing national security problems in detriment of protection of natural areas  [@Andrews2014]. But RPAS have also been generally welcomed in several developing countries in Asia, where an array of related programs are being carried out [@nugraha_urgency_2016]. [@Global2017] is a relative accurate database informing RPAS regulations by country. The uncertainty of the users along the world has promoted the development of associations in order to advise on the legal aspects to be taken into account during the operation, with the International Association for Unmanned Vehicle Systems (AUVSI) [@AUVSI2017]  being the largest nonprofit organization in the world dedicated to advancing the community of  unmanned aerial vehicles users. Also, a relative up-to-date database has been published online where users can consult the regulatory framework of RPAS by country [@Global2017].

Most of the sources analyzed focus on local-scale conservation projects and feasibility studies of RPAS monitoring distribution and abundance of wildlife populations. Literature begins to be equally prolific in mapping activities in terrestrial and aquatic ecosystems, a niche until recently entirely occupied by aerial and space platforms for environmental remote sensing. Despite the low number of scientific articles addressing the use of RPAS in the control and surveillance of natural protected areas, it is still one of the issues that more social debate generates and it is not strange to find governmental initiatives or promoted by environmental organizations in the fight against poaching. From the economic point of view,  expenses derived from the operation with RPAS are hardly quantifiable. Not all studies consider the effort required for the development of technical and analytical skills of work teams. Computational requirements are demanding and certain phases of information processing requires the acquisition of software whose price is generally high. Also, operations with RPAS are not exempt from accidents, which has an negative impact on the budget originally planned. In addition, statistical and sampling methods approaching the analysis of data are mostly in its infancy and further research should be accompased to assess the overall performance of these methods.

## Wildlife Monitoring and Management

Both fixed-wing and rotor-wing RPAS might become very handy tools for conservation practitioners. Wildlife census campaigns, usually carried out by going in on foot, terrestrial vehicles or by vessel deployment in aquatic environments, could be supplemented or replaced by RPAS mapping and monitoring capabilities. As becoming easier to operate, there are sufficient grounds to encourage park rangers training in the use of RPAS, which are subject in many cases to time-consuming and often dangerous raids. If operated responsibly, it could be closer of being considered a non-invasive and reliable monitoring technique [@jewell_effect_2013-1]. From the technological point of view, "Follow-me" capabilities of RPAS constitute a promising advance in animal movement and remote sensing disciplines, by having high-resolution aerial imagery from places frequently visited or crossed by electronically tagged species.     

## Infrastructure and risk assessment

Wildlife risk assessment could benefit from RPAS by promoting their use for preventive purposes in conflicting areas where human factor is indirectly causing the killing of many species.  Relative low operational cost of RPAS make them an attractive alternative to manual inspection, which may foster such practices. Since the literature citing RPAS for such purposes is limited, we propose that they could serve to reduce fatalities by scheduling periodic flights monitoring facilities, roads crossing sensitive areas and coastal ecosystems where vessels strikes with aquatic species is frequent. RPAS might also help to persuade birds from approaching power lines, wind turbines and other potential hazards, just as it has been applied to keep airports safe.

## Monitoring and mapping of terrestrial and aquatic ecosystems

The integration of classical remote sensing elements developed during the last decades in the scope of RPAS opens new possibilities in the observation of environmental phenomena at local scale, complementing other systems of Earth observation. Protected area managers should be aware of the benefits of having information on demand. As requirements change, Information Technology (IT) departments in protected areas must be ready to integrate data into effective conservation strategies and better decision-making. The inclusion of RPAS in monitoring activities should be weighed against the major complexity of such systems.

## Law enforcement

The convenience of using RPAS in the fight against poaching and illegal fishing in protected areas faces important technical and legal constraints. First, the reviewed literature mention the need to design more efficient live vision systems. Low autonomy of  RPAS is especially critical in large natural parks, limiting the area under surveillance. Issues concerning atmospheric conditions have not yet been completely resolved. [@banzi_sensor_2014] proposed a sensor based economical feasible anti-poaching alternative, arguing that RPAS fulfilling suitable specifications are costly, especially in developing countries. However, as technology becomes more accessible, it is expected that main barriers will appear in the legislative and social sphere. In some countries it is forbidden to fly beyond the visual range of the operator, limiting the effectiveness of the inspection in real time. RPAS applied to surveillance of protected areas is also questioned arguing human right breaching [@banzi_sensor_2014]. Some detractors are skeptical about the ability of RPAS to persuade offenders, who in many cases face situations of greatest need. Probably the success of such initiatives requires a greater consensus among the parties involved and the development of strategies that seek to solve the causes of poaching. Surveillance  of illegal logging activities  or bonfires detection in unauthorized areas have great potential and may be convenient to implement.

## Ecotourism

A permissive regularization of  RPAS in ecotourism activities in natural parks could lead to unpredictable situations. On the one hand, the constant presence of sources of noise coming from propellers and engines, the sensation of invasion or lack of privacy, security issues and the visual impact of RPAS on the landscape could negatively affect the tourist experience. It is well known that the consequences triggered by RPAS disturbing wildlife have led to the ban on flying them in national parks in the United States and other protected areas of the world. As result of potential enviromental impact due to the use of RPAS by tourist in Antarctica, [@leary_drones_2017] reported the partial prohibition of recreational RPAS in coastal areas as part of a more extensive regulation promoted by stakeholders. Such regulation looks reasonable and could be the way forward for other protected areas to adapt the allowed activities with RPAS. It seems obvious to deduce that in the hands of non-skilled operators, the risk of accidents and losses would increase. This may lead to the aforementioned wildlife disturbance, but they also pose a risk for contamination of water supplies or the triggering of fires in sensitive areas due to the presence of flammable and toxic components, fueling the low popularity of RPAS to the detriment of the benefits they bring. It does not appear that feasibility studies or opinion polls have been published that respond to the issues raised and to the ethical and legal implications derived from their use. Even when the leisure possibilities are wide and recognized, it would be advisable to be cautious in the face of the demand of the ecotourism industry to incorporate  RPAS in their activities. 

## Impact of RPAS on wildlife and ecosystems

The review of the literature suggests that there are still certain niches that need more attention from the research community. The ethical implications of RPAS in wildlife studies have not yet been adequately weighed since most studies only marginally address the presence or absence of reactions in species in the vicinity of RPAS. Despite the greater degree of awareness reflected in a emergent set of guidelines [@hodgson_best_2016; @mulero-pazmany_unmanned_2017] , we consider that further trials aimed at quantifying physiological and behavioral changes targeting a broader group of wild species should be carried out.  The establishment of a best practices and recommendations manual could increase the chances of integrating the responsible use of RPAS in conservation and management activities in natural areas.  Moreover, some authors mentioned the lack of commercial operators with sufficient expertise to carry out such activities [@mcevoy_evaluation_2016].  Also, an optimal trade-off between  benefits and environmental costs should be pursued [@gremillet_robots_2012; @sepulveda_review_2010]. By designing quieter, non-polluting and safer components, the impact on wildlife and ecosystems could be reduced and its objective observation facilitated. Nonetheless we trust that, as far as further testing be done, RPAS has great potential to replace more invasive monitoring techniques, whose reliability is challenged by the potential to induce conditions of unacceptable stress in wildlife that could ultimately invalidate the results of the research [@jewell_effect_2013, @wilson_measuring_2006]. This should be consciously taken into account by the many actors involved  in protected areas activities when they are reluctant to allow RPAS to be an essential part of research and conservation activities.
  
# Environmental monitoring and decision support
  
Protected areas are subject to periodic environmental quality control procedures where RPAS could play a major role.  Also, RPAS are suitable to assist decision making where rapid response is crucial by offering valuable information at real time to handle natural and man-made disasters. Wildfires is a major concern in natural parks and is not rare that RPAS have been put forward to assist in prevention, fighting and evaluation phases. In most cases, such applications have operational requirements which eventually are costly. For instance,  sophisticated on-board instruments, gas powered engines for longer endurance and higher payloads or robotics arms and containers designed to assist sampling, hold cargo or deliver assistance. RPAS could leverage wildlife capture procedures by carrying dart guns for chemical immobilization where otherwise manual approaching free-ranging animals is considered inefficient or dangerous.


# Conclusions 

The consolidation of the RPAS as management and research tools in protected areas is closely linked to the technological development of the elements associated with the platform and to the establishment of measures that favorably regulate its use, increasing opportunities in the sector and stimulating innovation in priority conservation areas. There are continually improvements in navigation control and flight autonomy, while we are witnessing the progressive miniaturization and diversification of sensors along with advances in the field of artificial intelligence. This rapidly expanding confluence of factors encourages the emergence of new scenarios with ethical and legal implications. Most governments have reacted by setting constraints that could have a negative impact on the capacity to integrate RPAS into the civilian sphere, despite some progress in this regard. As a result, it is difficult to foresee the actions that each country will adopt from now on in an attempt to harmonize the contradictions presented by RPAS, reason why it is probable that the future of the RPAS in protected areas is conditioned fundamentally by political and social factors.

# References 
