% interactcadsample.tex mod using svm.latex.ms.texs
% v1.03 - April 2017
\documentclass[]{interact}
%\documentclass[a4paper]{article}
\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required
% 
\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}


\usepackage{rotating}
% \usepackage[]{geometry}
\usepackage{url}




\usepackage{layout}
\setlength{\headsep}{15pt}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{tabu}
\onehalfspacing
\usepackage{helvet}


\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[nottoc,notlof,notlot]{tocbibind} 

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{
    \setlength{\leftmargin}{10mm}
    \setlength{\rightmargin}{\leftmargin}%
  }
  \relax}
 {\endlist}


\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
}
\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Jesús Jiménez López (University) and Margarita Mulero-Pázmány (University)},
             pdfkeywords = {RPAS, UAV, drones, natural protected areas, conservation, biodiversity},  
            pdftitle={Contribution of RPAS in research and conservation in protected areas:
present and future},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls


% notice vspace after title
\title{Contribution of RPAS in research and conservation in protected areas:
present and future\vspace{0.25in}  }


\author{\Large Jesús Jiménez López\vspace{0.05in} \newline\normalsize\emph{University }   \and \Large Margarita Mulero-Pázmány\vspace{0.05in} \newline\normalsize\emph{University }  }

\date{}


\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%555

\begin{document}

\articletype{ARTICLE TEMPLATE}

% \pagenumbering{arabic}% resets `page` counter to 1 
%
\author{
\name{A.~N. Author\textsuperscript{a}\thanks{CONTACT L.~Jimenez. Jesus. Email: foo@foo.com} and Margarita Múlero-Pázmani\textsuperscript{b}}
\affil{\textsuperscript{a}Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK; \textsuperscript{b}Institut f\"{u}rDepartment of Evolutionary Ecology, Doñana Biological Station, CSIC\"{a}t,  Seville, Spain}
}
% \author{\Large Jesús Jiménez López\vspace{0.05in} \newline\normalsize\emph{University}  \and \Large Margarita Mulero-Pázmány\vspace{0.05in} \newline\normalsize\emph{University} }
\title{Contribution of RPAS in research and conservation in protected areas:
present and future}
\maketitle  % title \par 
  




  \fontsize{9}{10}{\textbf{ABSTRACT}}: \par
  \begingroup
  \leftskip1em
  \rightskip\leftskip
  \noindent{\fontsize{9}{10}\selectfont In the last few decades, we have witnessed a growing interest in
projects aimed to evaluate the feasibility of RPAS for conservation
purposes. So far, RPAS have been proposed for a variety of research and
management activities including environmental and wildlife monitoring or
law enforcement. It remains to be seen how well RPAS complement
conservation actions in protected areas, since beyond technical, ethical
and legal barriers that compromise their effectiveness, it is a common
source of discussion that bias in research efforts is often of limited
utility in addressing practical issues, particularly those claimed by
natural park managers. We carried out a systematic review of the
published literature in order to consider whether more attention should
be given to certain conservation activities where the contribution of
RPAS could improve the efficiency in the management of protected areas.
After discussing results, we expose main drawbacks and identify those
areas that requires further research investment, highlighting some
trends and opportunities that apparently have not yet been adequately
exploited.}
  \rightskip5em
  \par
  \endgroup
  \vspace{5mm}
  


    
 \fontsize{9}{10}{\textbf{KEYWORDS}}: \par
  \begingroup
  \leftskip1em
  \rightskip\leftskip
  \noindent{\fontsize{9}{10}\selectfont RPAS, UAV, drones, natural protected areas, conservation, biodiversity}
  \rightskip5em
  \par
  \endgroup
  \vspace{5mm}


% 

\vskip 6.5pt

\noindent  \section{Introduction}\label{introduction}

As defined by UICN, ``a protected area is a clearly defined geographical
space, recognized, dedicated and managed, through legal or other
effective means, to achieve the long term conservation of nature with
associated ecosystem services and cultural values'' (Dudley 2008).
Protected areas have been declared under different reasons and
circumstances but there is a general consensus on its importance in
safeguarding biodiversity, contribute to human well-being and ensure
persistece of the natural heritage for the enjoyment of future
generations. Despite such praiseworthy intentions, the reality faced by
protected areas is subject to a wide variety of requirements and
unforeseen challenges requiring rapid and effective solutions.

To ensure its ability to meet both long-term and short-term conservation
goals, protected areas have benefited from a extensive regulatory
framework, to a certain extent implemented through the management plans.
These plans aimed to establish the basic guidelines to protect the
natural and socioeconomic values by which protected areas were declared.
In the field of competences derived from such instruments, available
human and material resources have mainly been allocated to regulate
tourism and recreational activities, monitoring law enforcement
including various forms of illegal resource extraction, wildfire
prevention and fighting, wildlife monitoring campaigns to maintain
up-to-date fauna inventories, environmental assesment or actions aimed
at strengthen environmental education and research programs. Those
activities have benefit from a wide range of technological advances,
including remote sensors, field monitoring stations, manned surveys,
camera traps, wildlife tracking devices or computational tools. More
recently, applications of remotely piloted aircraft systems (RPAS, also
known as unmanned aerial systems, UAS, drones) have been the subject of
a growing interest in both the civilian sphere and scientific community.
Not surprisingly, there have been a significant amount of research
projects facing conservation issues using RPAS, especially during the
last decade (Linchant et al. 2015; Chabot and Bird 2015; Christie et al.
2016). As a consequence, RPAS have been suggested as an appropriate
complement for conservation activities (Koh and Wich 2012; Rodríguez et
al. 2012; Chabot and Bird 2015, Wulder et al. (2004), Zahawi et al.
(2015)).

However, to date, it has not been adequately considered whether research
efforts using RPAS meet the demands of conservation practitioners. The
fact that the RPAS has received increasing attention in studies
addressing its feasibility to approach ecological issues and practical
conservation problems does not necessarily respond to the priorities
raised. In addition, budgetary constraints are common in protected area
management and available resources are generally insufficient for
optimal compliance. While park managers are responsible for acquiring as
much information as possible to ensure that such potential source of
investment will have a positive effect on the management efficiency and
quality of services provided, the commitment of the research community
is to provide the appropriate responses to facilitate better
decision-making. We note that wildlife and habitat monitoring has
received considerable attention. However, the scarcity of lines of
research in other areas of equal importance is an important gap to
consider when attempting to demonstrate the true potential of RPAS,
encouraging its formal application in protected area management tasks.

However, to date, it has not been adequately considered whether research
efforts using RPAS meet the demands of conservation practitioners. As
noted, wildlife and habitat monitoring have received major emphasis.
Consequently, this effort does not necessarily respond to practical
conservation priorities. In addition, budgetary constraints are common
in protected area management and available resources are generally
insufficient for optimal compliance. Park managers must have sufficient
guarantees to ensure that such potential source of investment will have
a positive effect on the management efficiency and quality of services
provided. The commitment of the research community is to provide the
appropriate responses to facilitate better decision-making. Fostering
RPAS in protected areas requires cover this important gap to consider
when attempting to demonstrate the true potential of RPAS, encouraging
its formal application in protected area management tasks.

, while there are important gaps to consider when uncovering the true
potential of RPAS for conservation tasks.

There is an uneven amount of studies facing

\section{Methods}\label{methods}

To achieve the proposed objectives, a systematic bibliographical review
(see PRISMA Flowchart) of scientific articles, gray literature,
postgraduate theses, websites and specialized journals was carried out,
following a similar line to other related studies (Linchant et al. 2015;
Christie et al. 2016; Mulero-Pázmány et al. 2017). Last reference
revised was published on June, 2017. The main tools for selecting
bibliography include Google Schoolar, Research Gate and Mendeley
Desktop, while the use of Internet search engines include other
references outside the scientific scope. Key search criteria, primarily
in English, include RPAS in their various meanings and acronyms,
reflecting the varied terminology used: ``remotely piloted aerial system
(RPAS)'', ``drone'', ``unmaned aircraft system (UAS)'' ``unmaned
aircraft'', ``model aircraft'', ``radio control aircraft'' , ``unmaned
aerial vehicle (UAV)'' along with several terms referring to common
activities carried out in protected areas, from the general to the
particular: ``conservation'', ``protected areas'', ``ecology''
``ecosystem'', ``habitat'', ``vegetation'', ``forest'', ``wetland'',
``reforestation'', ``monitoring'', ``survey'', ``sampling'',
``inventory'', ``wildlife'', ``fauna'', ``bird'', ``mammal'',
``fish,''amphibian``,''reptile``,''wildfire``,''remote
sensing``,''tourism``,''ecotourism``,''law
enforcement``,''poaching``,''logging``,''risk management``,''pollution"
and so on. For example, a restricted search was carried out with a
logical statement as follows: ``conservation drone'' OR ``conservation
UAS'' OR ``conservation UAV'' OR ``conservation unmanned aircraft
system'' OR ``conservation unmanned aircraft'' OR ``conservation RPAS''
OR ``conservation radio control aircraft'', both in singular and plural
forms, as there are limits in the number of words in a single search
statement. That counts for more than 80 posible combinations. We also
included references from above mentioned review articles and those not
indexed in research results but relevant for the purpose of this study.

From the total number of publications (x), we excluded duplicated
results or those with no relevance for conservation, based on activities
included in protected area management plans worldwide. The selected
publications (x) were categorized according to the main purpose of the
study. These categories include: ``wildlife monitoring and management'',
for those feasibility studies facing alternative population surveys and
fauna tracking methods, together with advances in software development
focused on automatic animal counting and detection; ``monitoring and
mapping of terrestrial and aquatic ecosystems'', for vegetation surveys,
ecosystem dynamics studies and habitat characterization; ``Law
enforcement'' encompasses poaching, illegal logging and fishing
activities; ``ecotourism'' refers mainly to recreational activities and
visitors management; ``risk management, enviromental assesment and
decision support'' encompass a wide variaty of duties, including
pollution sampling, disaster management. Finally, legal constraints and
actions to minimize impact on fauna are also considered, as both shape
the feasibility of RPAS to approach conservation and environmental
issues.

The collected information is presented in tabular format, identifying
where the study was conducted, the expected accomplishments and
technical specifications of the aerial platform. After expose main
results, gaps are identified and possible scenarios for implementing
RPAS as essential tools to help achieve conservation goals in protected
areas are discussed.

\section{Results}\label{results}

\subsection{Wildlife Monitoring and
Management}\label{wildlife-monitoring-and-management}

Wildlife science generally refers to the study, monitoring, and
management of wild mammals, birds, reptiles, and amphibians (and
occasionally fish), as well as their habitats.

Several studies focused on develop survey methods aimed at locating,
counting and characterizing wildlife communities and habitats, ranging
from large and medium size terrestrial mammals (Lancia et al. 2005; Jain
2013, Barasona et al. (2014)), birds (Wilson, Barr, and Zagorski 2017,
Hodgson et al. (2016),Christie et al. (2016) Sardà-Palomera et al.
(2012), Chabot and Bird (2012)), but also species relying on coastal and
marine ecosystems (Colefax, Butcher, and Kelaher 2017; Hodgson, Peel,
and Kelly 2017; {\textbf{???}}; Koski et al. 2015, Dulava, Bean, and
Richmond (2015), Durban et al. (2015), Koski et al. (2009)). RPAS have
as well been applied as a means of inspecting and assess breeding and
nesting areas at inaccessible sites in several species (Szantoi et al.
2017; Wich et al. 2016, Puttock et al. (2015); Andel et al. 2015;
Weissensteiner, Poelstra, and Wolf 2015 ) or as a complement for
wildlife telemetry tracking methods {[}Christie et al. (2016); Bayram et
al. (2016); Mulero-Pázmány et al. (2015); Körner et al. (2010); Cliff et
al. (2015); Leonardo et al. (2013); Soriano, Caballero, and Ollero
(2009);{]}. Given the large amount of information generated, it is not
surprising that computer vision software have been developed to handle
the automatic detection, recognition and counting of individuals
captured in scenes acquired by visible and thermal-infrared sensors,
replacing otherwise time-consuming manual tasks (Andrew and Shephard
2017; Chabot and Francis 2016; Gonzalez et al. 2016, Lhoest et al.
(2015); Gemert et al. 2015; Christiansen et al. 2014; Martin et al.
2012; Abd-Elrahman, Pearlstine, and Percival 2005).

\newpage

\blandscape

\tiny

\begin{longtable}[]{@{}llllllll@{}}
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Mulero-Pázmány et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Telemetry/RPAS SDM comparative study\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Spain\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Doñana National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
\emph{Bos taurus}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Easy Fly plane, Ikarus autopilot, Eagletree GPS logger\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Panasonic Lumix LX-3 11MP\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 6500\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Hodgson, Peel, and Kelly 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Comparative survey RPAS/land based observation; abundance
estimation\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
North Stradbroke Island\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
humback whales\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: ScanEagle\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Nikon D90 12MP, Standard Definition Electro-Optical Camera\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Hodgson, Kelly, and Peel 2013)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Dugongs detection\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Shark Bay Marine Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Dugong\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: ScanEagle\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Nikon D90 12MP\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Wilson, Barr, and Zagorski 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Bioacustic monitoring\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
State Game Lands\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Birds\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom 2\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
ZOOM H1 Handy Recorder\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Bayram et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
VHF collars tracking\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Bears (Ursus)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI F550\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Telonics MOD-500 VHF, Uniden handheld scanner\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Christie et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Abundance estimation\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Aleutian Islands\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Steller Sea Lion (Eumetopias jubatus)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: APH-22\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 25.000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Christie et al. 2016\})\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Abundace estimation\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Monte Vista National Wildlife Refuge\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Grus canadensis (sandhill cranes)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Raven RQ- 11A\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 400\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Wich et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Sumatran orangutan nest detection\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Skywalker 2013\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon S100\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Andel et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Chimpanzee nest detection\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Africa\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Loango National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Chimpanzee (Pan troglodytes)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Maja\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon Powershot SX230 HS\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 5000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Koski et al. 2009)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Marine mammals monitoring\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Admiralty Bay\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Marine mammals\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: ScanEagle\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
NTSC Video Camera\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Andrew and Shephard 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Semi-automated image processing tools to detect and map sea eagle
nests\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Houtman Abrolhos Islands\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
White-bellied sea eagle (Haliaeetus leucogaster)\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Longmore et al. 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Software development to help detect animals in thermal images\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
UK\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Arrowe Brook Farm Wirral\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Wildlife\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: 3DR robotics Y6\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
FLIR, Tau 2 LWIR Thermal Imaging Camera Core\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Martin et al. 2012)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Estimate the distribution of organisms using statistical models\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Manatee (\emph{Trichechus manatus latirostris})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Nova 2.1\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Olympus H E-420\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
{[}(Chabot and Bird 2012){]}\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Waterbirds surveys\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Canada\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Ste-Anne-de-Bellevue, Baie-du-Feb- vre\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Canada Geese (\emph{Branta canadensis}), Snow Geese (\emph{Chen caerule-
scens})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing; CropCam\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Pentax Optio A20 10MP\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 11000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Koski et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Reidentification of photographed whales\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Canada\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Igloolik Island\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Bowhead whales (\emph{Balaena mysticetus})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: TD100E\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
GoPro, Nikon D800 50 mm f1.2 Nikon lens\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Durban et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Photogrammetry studies\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Canada\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Vancouver Island\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Killer whales (\emph{Orcinus orca})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: APH-22\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Olympus E-PM2 Olympus M.Zuiko 25 mm F1.8 lens\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Lhoest et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Develop algorithm for automatic count of hippos\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Congo\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Garamba National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Common hippopotamus (\emph{Hippopotamus amphibius L.})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Falcon\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Tamarisk 640\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Goebel et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Estimating abundance and size of Antarctic predators\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Antarctica\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Cape Shirreff, Livingston Island, South Shetland Islands\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Antarctic fur seal (\emph{Arctocephalus gazella}), Penguin
(\emph{Spheniscidae}), Leopard seal (\emph{Hydrurga leptonyx})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: md4-1000, APQ-18, APH-22\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Sony NEX-5, Canon EOS-M,Olympus E-P1\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Dulava, Bean, and Richmond 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Waterbirds surveys\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Ruby Lake NWR, Tomales Bay\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
mallard (\emph{Anas platyrhynchos}), northern shoveler (\emph{Anas
clypeata}), gadwall (\emph{Anas strepera}), blue-winged teal (\emph{Anas
discors}), northern pintail (\emph{Anas acuta}), American wigeon
(\emph{Anas americana}), canvasback (\emph{Aythya valisineria}), scaup
(\emph{Aythya marila/ Aythya affinis}), surf scoter (\emph{Melanitta
perspicillata}), American coot (\emph{Fulica americana})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Honeywell RQ-16 T-Hawk , Fixed-wing: AeroVironment
RQ-11A\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon PowerShot SX230, SX260 12.1MP, GoPro Hero3\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Cliff et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Online Localization of Radio-Tagged Wildlife\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Noisy miner (M\emph{anorina Melanocephala})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Falcon 8\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Radiometrix LMR1 receiver, Digi XTend radio modem\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
({\textbf{???}})\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Salmon spawning surveys\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Lower Snake River\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Chinook salmon (\emph{Oncorhynchus tshawytscha})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Aeryon Scout\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
GoPro Hero3\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Barasona et al. 2014)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Epidemiological study\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Spain\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Doñana National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Red deer (\emph{Cervus elaphus}), Fallow deer (\emph{Dama dama}), Cattle
(\emph{Bos taurus})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Easy Fly plane\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Panasonic Lumix LX-3 11MP\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Jain 2013)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Elephant surveys\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Burkina Faso\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Nazinga Game Ranch\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Elephant (\emph{Loxodonta africana})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Gatewing x100\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Ricoh GR III\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Rodríguez et al. 2012)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
UAS and GPS Data Loggers\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Spain\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Doñana National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Lesser kestrel (\emph{Falco naumanni})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: ST-model Easy Fly\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Panasonic Lumix LX-3\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Sardà-Palomera et al. 2012)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Fine-scale bird monitoring\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Spain\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Estany d'Ivars i Vila-Sana\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Black-headed (\emph{Gull Chroicocephalus ridibun})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
fixed-wing: Multiplex Twin Star II\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Panasonic Lumix FT-1 12MP 28--128 mm\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 1645\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Hodgson et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Ccolony nesting birds surveys\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Ashmore Reef Commonwealth Marine Reserve, Adele Island, Macquarie
Island\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Frigatebirds (\emph{Fregata ariel}), Crested Terns (\emph{Talasseus
bergii}), Royal Penguins (\emph{Eudyptes schlegeli})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: X8 3D Robotics, Fixed-wing: FX79\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon EOS M 40mm\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 5100\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

{[}@-{]} \textbar{} - \textbar{} - \textbar{} - \textbar{} - \textbar{}
- \textbar{} \textbar{} - \textbar{}

\begin{longtable}[]{@{}llllllll@{}}
\caption{WILDLIFE MONITORING AND MANAGEMENT}\tabularnewline
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Perroy, Sullivan, and Stephenson 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Tropical invasive plants\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Pahoa, Hawai\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
\emph{Miconia calvescens}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJ Inspire-1\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
DJI FC350 camera\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Szantoi et al. 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Habitat Mapping\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Indonesia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Gunung Leuser National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Orangutan (\emph{Pongo abelii})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Skywalker\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon S100\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 4000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Casella et al. 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Coral reef mapping\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
French Polynesia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Tiahura,; Moorea\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Coral reef\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom 2\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Modified GoPro HERO4\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 1678\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Casella et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Monitoring coastal erosion dynamics in shorelines\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
French Polynesia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Tiahura; Moorea\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Coral reef\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Mikrokopter Okto XL\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon G11\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 7500\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Müllerová et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Monitoring plant invasion\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Exotic species\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: VUT 712 713 720\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon S100\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Ventura et al. 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Marine fish nursery areas mapping\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Italy\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Giglio Island\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Marine fish nursery areas\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: homemade prototype\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Mobius HD, GoPro HERO3 Black Edition\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 100\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Ivošević et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Habitat monitoring and modeling in restricted areas; RPAS performance
test \& South Korea \& Chiaksan National Park;Taeanhaean National
Park\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
South Korea\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Chiaksan National Park;Taeanhaean National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom 2 Vision+\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Full HD videos 1080p/30fps and 720p/60fps\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Lisein et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Discrimination of deciduous species; Forest inventory\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Belgium\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Grand-Leez\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
English oak, birches, sycamore maple ,common ash and poplars\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wings: Gatewing X100\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Ricoh GR2 GR3 GR4 10 megapixels CCD\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Puttock et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Characterization of ecosystems affected by beaver activity\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
UK\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Devon Beaver Project site\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Eurasian beaver (\emph{Castor fiber})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: 3D Robotics Y6\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
anon ELPH 520 HS\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Zahawi et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Characterization of tropical forest structure for restoration
actions\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Costa Rica\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Devon Beaver Project site\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Several species\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: 3D Robotics Y6\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon S100\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 1500\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Bustamante 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Forest monitoring\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Brasil\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Riverine Forests (Permanent Protected Areas), Rio de Janeiro, Barrãcao
do Mendes, Santa Cruz and São Lorenço\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Riverbank forests\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom Vision 2S\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
RGB digital camera 14MP\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 9700\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Gini et al. 2012)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
3D modeling and classification of tree species\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Italy\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Parco Adda Nord\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Several species\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Microdrones TM MD4-200\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
RGB CCD 12 megapixels Pentax Optio A40; modified NIR Sigma DP1 with a
Foveon X3 sensor\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Miyamoto et al. 2004)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Classification of species in wetlands\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Japan\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Kushiro Wetlands\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Several species\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Helium balloon\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
NIKON F-801, NIKKOR 28 mm f/2.8\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 1600\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Casella et al. 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Mapping coral reefs\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor.wing: DJI Phantom 2\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
GoPro HERO4\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}llllllll@{}}
\caption{MONITORING OF TERRESTRIAL AND AQUATIC
ECOSYSTEMS}\tabularnewline
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Lobermeier et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Mitigate the risk of collision by installing markers on electrical
lines\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Birds\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Mikrokopter Hexa XL\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
KX 171 Microcam\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Margarita Mulero-Pázmány 2014\})\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Bird risk hazards in power lines\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Spain\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Doñana National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Birds\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
ixed-wing: Easy fly St-330\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
GoPro HERO2 11MP;Panasonic LX3 11MP\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 8863\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Mulero-Pázmány et al. 2014)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Anti-poaching\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Africa\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
KwaZulu-Nata\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Black rhinocero (\emph{Diceros bicornis}), white rhinocero
(\emph{Ceratotherium simum})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Easy Fly St-330\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Panasonic Lumix LX-3 11 MP, GoPro Hero2; Thermoteknix Micro CAM
microbolometer\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 15700\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Hansen 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Visitors Surveillance\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Sweden\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Sweden \& Kosterhavet National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Humans\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Sabella et al. 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Visitors Surveillance\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Italy\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
R.N.O. Oasi faunistica di Vendicari\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Humans\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom 3\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(King 2014)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
RPAS applications in ecotourism activities\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Sweeden\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Sweeden \& Kosterhavet National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Humans\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Vas et al. 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
RPAS impact\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
France\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Zoo du Lunaret, Cros Martin Natural Area\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
\emph{Anas platyrhyncho}, \emph{Phoenicopterus roseus}, \emph{Tringa
nebularia}\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Phantom\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
GoPro HERO3\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Weissensteiner, Poelstra, and Wolf 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
RPAS Impact\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Sweeden\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Hooded crow (\emph{Corvus corone cornix})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom 2 Vision\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 1000\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}llllllll@{}}
\caption{INFRASTRUCTURES AND RISK ASSESSMENT, ECOTOURISM, IMPACT ON
WILDLIFE AND ECOSYSTEMS}\tabularnewline
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Study\strut
\end{minipage} & \begin{minipage}[b]{0.18\columnwidth}\raggedright\strut
Aims\strut
\end{minipage} & \begin{minipage}[b]{0.03\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Place\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
Target\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
RPAS platform\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Payload\strut
\end{minipage} & \begin{minipage}[b]{0.01\columnwidth}\raggedright\strut
Costs\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Zang et al. 2012)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Pollution monitoring\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
China\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Several cities\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Rivers\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon 50D, ACD multispectral camera\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Cornell, Herman, and Ontiveros 2016)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Water sampling\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Lake Ontari0\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: DJI Phantom 3\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
50mL Falcon tube\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(McCaldin, Johnston, and Rieker 2015)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Aerial baiting\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Christmas Island\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Cat (\emph{Felis catus})\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: V-TOL Hornet I-II\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Canon S100; Drop mechanism with HD Video Recorder\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Fornace et al. 2014)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Spatial epidemiology\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Malaysia / Philippines\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Sabah / Palawan\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
?\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Sensefly eBee\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
16mp\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
\$ 25000\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Van Tilburg 2017)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Search and Rescue (SAR)\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Columbia Gorge National Scenic Area\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Humans\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Phantom 3, SAR Bot, Inspire 1\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
DJI 12MP; VUE PRO 640 thermal imager\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Schwarzbach et al. 2014)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Water sampling\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
Spain\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Doñana National Park\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Freshwater ecosystems\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Rotor-wing: Helicopter\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Water sampling mechanism\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
(Schmale, Dingus, and Reinholtz 2008)\strut
\end{minipage} & \begin{minipage}[t]{0.18\columnwidth}\raggedright\strut
Aerobiological sampling\strut
\end{minipage} & \begin{minipage}[t]{0.03\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Virginia Tech's Kentland Farm\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
Prokaryotic and eukaryotic microorganisms\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
Fixed-wing: Senior Telemaster\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Aerobiological sampling devices\strut
\end{minipage} & \begin{minipage}[t]{0.01\columnwidth}\raggedright\strut
?\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Table: ENVIRONMENTAL MONITORING AND DECISION SUPPORT

\elandscape
\normalsize

\subsection{Monitoring and mapping of terrestrial and aquatic
ecosystems}\label{monitoring-and-mapping-of-terrestrial-and-aquatic-ecosystems}

So far RPAS have been used to classify vegetation cover in wetlands
(USGS 2014), monitoring the spread and detection rate of invasive
species (Müllerová et al. 2016, Zaman, Jensen, and McKee (2011), Perroy,
Sullivan, and Stephenson (2017)), the characterization of forest stands
(Gini et al. 2012), including the evaluation of the effectiveness of
restoration actions (Zahawi et al. 2015) or multitemporal analysis of
the spectral response to phenological variations in different species of
deciduous trees (Lisein et al. 2015). Recently, shallow coastal habitats
were also mapped using consumer grade RPAS (Casella et al. 2017, Ventura
et al. (2016)), but also to monitor erosion dynamics in shorelines.
(Casella et al. 2016). This momentum has been especially notable with
the parallel development of affordable multispectral and hyperspectral
sensors adapted to small aircraft (Bareth et al. 2015).

\subsection{Infrastructure and risk
assessment}\label{infrastructure-and-risk-assessment}

Other research projects highlight the convenience of RPAS in assessing
the risk that linear electrical infraestructures posed for birds
(Margarita Mulero-Pázmány 2014; Lobermeier et al. 2015) or identify
nests on the ground at risk of destruction by harvesting (Mulero-Pázmány
Margarita 2011).

\subsection{Law enforcement}\label{law-enforcement}

RPAS have also relevance in the control and surveillance of protected
areas, documented through different experiences and considerations
focused mainly on poaching and illegal fishing activities
(Mulero-Pázmány et al. 2014, Franco et al. (2016), Olivares-Mendez et
al. (2014)) but also other less contentious illegal activities (Sabella
et al. 2017). (Duffy 2014) analyzed the consequences of the
militarization of conservation practices as an increasing trend in
natural protected areas around the world and illustrates the use of RPAS
through several examples.

\subsection{Ecotourism}\label{ecotourism}

The high degree of diversification offered by RPAS in the ecotourism
industry is summarized in a recent article, which shows possible
recreational activities, business opportunities, search and rescue
operations, mapping and formulas for granting RPAS flight permits in
designated areas (King 2014). Within the still scarce literature,
(Hansen 2016) values the effectiveness of RPAS in monitoring visitors in
marine and coastal areas, in combination with other innovative
solutions. According to the author, RPAS would theoretically allow to
operate under different environmental conditions, improving the level of
detail and offering a continuous coverage in the flow and behavior of
visitors, as opposed to other techniques of habitual use like the manual
observation or the installation of networks of surveillance cameras.

\section{Environmental management and decision
support}\label{environmental-management-and-decision-support}

Planning in protected areas is reflected through a variety of management
programs that are difficult to fit into some of the previously discussed
categories. An extensive report based on Christmas Island proposed
dropping poisoned baits from RPAS to eradicate feral cats disturbing
threatened native species (McCaldin, Johnston, and Rieker 2015).
(Cornell, Herman, and Ontiveros 2016) obtained ground truth data by
adapting RPAS to take water samples for comparison with hyperspectral
measurements of Landsat 8 Operational Land Imager (OLI). (Zang et al.
2012) identified several pollution agents in riparian areas using RPAS
imagery. (Schwarzbach et al. 2014) goes further away by performing
several aerial water sampling methods using an unmaned helicopter to
monitor water pollution, while (Schmale, Dingus, and Reinholtz 2008)
collected a broad spectrum of both prokaryotic and eukaryotic
microorganisms using a fixed-wing aircraft equiped with a custom made
aerial sampling device. (Fornace et al. 2014) considered mapping
enviromental risk factors for predicting zoonotic diseases as part of a
extensive epidemiological study carried out in Philippines and so on.
Literature citing RPAS for search and rescue activities is profuse and
an in-depth revision is beyond the scope of this article, but a recent
publication ilustrate several examples where RPAS were succesfully
operated to assist rescue teams (Van Tilburg 2017). A google scholar
search sorted by relevance using disaster management and drones keywords
throws at first place a complete report describing a complex framework
for decision support using RPAS (Maza et al. 2011).

\subsection{Impact of RPAS on wildlife and
ecosystems}\label{impact-of-rpas-on-wildlife-and-ecosystems}

Animal welfare in wildlife management practices and ecological research
is a sensitive issue from which ethical issues arise (Wilson and McMahon
2006). Not surprisingly, RPAS are not exempt of discussion and several
trials measure the disturbance effects of RPAS on birds (Vas et al.
2015, McEvoy, Hall, and McDonald (2016),BORRELLE and FLETCHER (2017),
Scobie and Hugenholtz (2016), Weissensteiner, Poelstra, and Wolf (2015))
and mammals (Ditmer et al. 2015, Pomeroy, O'Connor, and Davies (2015)),
while other studies marginally inform observed behavioural patterns
(Jain 2013, Mulero-Pázmány et al. (2015)). Finally,a code of good
practice and recommendations is build upon the learned experiences.

\section{Discussion}\label{discussion}

\subsection{Legal barriers}\label{legal-barriers}

RPAS operations faces important legal barriers that undermine the true
potential in the civilian sphere (Stöcker et al. 2017). An overly
restrictive regulatory framework could limit the possibilities of use of
the RPAS in the field of conservation, which makes clear the urgent need
to harmonize legislation . In the United States and in most of the
European countries consulted, interim legislation has been adopted
which, to a certain extent, equates the management of RPAS with that of
traditional aircraft. In general terms, the situation in Latin America
is uneven, however there is a general tendency to develop specific laws
to cope with the rise of the RPAS in both the civil and military sectors
(America 2017). Africa is one of the continents where the impact of RPAS
in conservation has had greater repercussions. However, in the opinion
of some conservationists, their use have not been without problems,
resulting in governments that have totally or partially prohibited drone
operations, arguing national security problems in detriment of
protection of natural areas (Andrews 2014). But RPAS have also been
generally welcomed in several developing countries in Asia, where an
array of related programs are being carried out (Nugraha, Jeyakodi, and
Mahem 2016). (Consulting 2017) is a relative accurate database informing
RPAS regulations by country. The uncertainty of the users along the
world has promoted the development of associations in order to advise on
the legal aspects to be taken into account during the operation, with
the International Association for Unmanned Vehicle Systems (AUVSI)
(UAVSI 2017) being the largest nonprofit organization in the world
dedicated to advancing the community of unmanned aerial vehicles users.
Also, a relative up-to-date database has been published online where
users can consult the regulatory framework of RPAS by country
(Consulting 2017).

Most of the sources analyzed focus on local-scale conservation projects
and feasibility studies of RPAS monitoring distribution and abundance of
wildlife populations. Literature begins to be equally prolific in
mapping activities in terrestrial and aquatic ecosystems, a niche until
recently entirely occupied by aerial and space platforms for
environmental remote sensing. Despite the low number of scientific
articles addressing the use of RPAS in the control and surveillance of
natural protected areas, it is still one of the issues that more social
debate generates and it is not strange to find governmental initiatives
or promoted by environmental organizations in the fight against
poaching. From the economic point of view, expenses derived from the
operation with RPAS are hardly quantifiable. Not all studies consider
the effort required for the development of technical and analytical
skills of work teams. Computational requirements are demanding and
certain phases of information processing requires the acquisition of
software whose price is generally high. Also, operations with RPAS are
not exempt from accidents, which has an negative impact on the budget
originally planned. In addition, statistical and sampling methods
approaching the analysis of data are mostly in its infancy and further
research should be accompased to assess the overall performance of these
methods.

\subsection{Wildlife Monitoring and
Management}\label{wildlife-monitoring-and-management-1}

Both fixed-wing and rotor-wing RPAS might become very handy tools for
conservation practitioners. Wildlife census campaigns, usually carried
out by going in on foot, terrestrial vehicles or by vessel deployment in
aquatic environments, could be supplemented or replaced by RPAS mapping
and monitoring capabilities. As becoming easier to operate, there are
sufficient grounds to encourage park rangers training in the use of
RPAS, which are subject in many cases to time-consuming and often
dangerous raids. If operated responsibly, it could be closer of being
considered a non-invasive and reliable monitoring technique (Jewell
2013a). From the technological point of view, ``Follow-me'' capabilities
of RPAS constitute a promising advance in animal movement and remote
sensing disciplines, by having high-resolution aerial imagery from
places frequently visited or crossed by electronically tagged species.

\subsection{Infrastructure and risk
assessment}\label{infrastructure-and-risk-assessment-1}

Wildlife risk assessment could benefit from RPAS by promoting their use
for preventive purposes in conflicting areas where human factor is
indirectly causing the killing of many species. Relative low operational
cost of RPAS make them an attractive alternative to manual inspection,
which may foster such practices. Since the literature citing RPAS for
such purposes is limited, we propose that they could serve to reduce
fatalities by scheduling periodic flights monitoring facilities, roads
crossing sensitive areas and coastal ecosystems where vessels strikes
with aquatic species is frequent. RPAS might also help to persuade birds
from approaching power lines, wind turbines and other potential hazards,
just as it has been applied to keep airports safe.

\subsection{Monitoring and mapping of terrestrial and aquatic
ecosystems}\label{monitoring-and-mapping-of-terrestrial-and-aquatic-ecosystems-1}

The integration of classical remote sensing elements developed during
the last decades in the scope of RPAS opens new possibilities in the
observation of environmental phenomena at local scale, complementing
other systems of Earth observation. Protected area managers should be
aware of the benefits of having information on demand. As requirements
change, Information Technology (IT) departments in protected areas must
be ready to integrate data into effective conservation strategies and
better decision-making. The inclusion of RPAS in monitoring activities
should be weighed against the major complexity of such systems.

\subsection{Law enforcement}\label{law-enforcement-1}

The convenience of using RPAS in the fight against poaching and illegal
fishing in protected areas faces important technical and legal
constraints. First, the reviewed literature mention the need to design
more efficient live vision systems. Low autonomy of RPAS is especially
critical in large natural parks, limiting the area under surveillance.
Issues concerning atmospheric conditions have not yet been completely
resolved. (Banzi 2014) proposed a sensor based economical feasible
anti-poaching alternative, arguing that RPAS fulfilling suitable
specifications are costly, especially in developing countries. However,
as technology becomes more accessible, it is expected that main barriers
will appear in the legislative and social sphere. In some countries it
is forbidden to fly beyond the visual range of the operator, limiting
the effectiveness of the inspection in real time. RPAS applied to
surveillance of protected areas is also questioned arguing human right
breaching (Banzi 2014). Some detractors are skeptical about the ability
of RPAS to persuade offenders, who in many cases face situations of
greatest need. Probably the success of such initiatives requires a
greater consensus among the parties involved and the development of
strategies that seek to solve the causes of poaching. Surveillance of
illegal logging activities or bonfires detection in unauthorized areas
have great potential and may be convenient to implement.

\subsection{Ecotourism}\label{ecotourism-1}

A permissive regularization of RPAS in ecotourism activities in natural
parks could lead to unpredictable situations. On the one hand, the
constant presence of sources of noise coming from propellers and
engines, the sensation of invasion or lack of privacy, security issues
and the visual impact of RPAS on the landscape could negatively affect
the tourist experience. It is well known that the consequences triggered
by RPAS disturbing wildlife have led to the ban on flying them in
national parks in the United States and other protected areas of the
world. As result of potential enviromental impact due to the use of RPAS
by tourist in Antarctica, (Leary 2017) reported the partial prohibition
of recreational RPAS in coastal areas as part of a more extensive
regulation promoted by stakeholders. Such regulation looks reasonable
and could be the way forward for other protected areas to adapt the
allowed activities with RPAS. It seems obvious to deduce that in the
hands of non-skilled operators, the risk of accidents and losses would
increase. This may lead to the aforementioned wildlife disturbance, but
they also pose a risk for contamination of water supplies or the
triggering of fires in sensitive areas due to the presence of flammable
and toxic components, fueling the low popularity of RPAS to the
detriment of the benefits they bring. It does not appear that
feasibility studies or opinion polls have been published that respond to
the issues raised and to the ethical and legal implications derived from
their use. Even when the leisure possibilities are wide and recognized,
it would be advisable to be cautious in the face of the demand of the
ecotourism industry to incorporate RPAS in their activities.

\subsection{Impact of RPAS on wildlife and
ecosystems}\label{impact-of-rpas-on-wildlife-and-ecosystems-1}

The review of the literature suggests that there are still certain
niches that need more attention from the research community. The ethical
implications of RPAS in wildlife studies have not yet been adequately
weighed since most studies only marginally address the presence or
absence of reactions in species in the vicinity of RPAS. Despite the
greater degree of awareness reflected in a emergent set of guidelines
(Hodgson and Koh 2016; Mulero-Pázmány et al. 2017) , we consider that
further trials aimed at quantifying physiological and behavioral changes
targeting a broader group of wild species should be carried out. The
establishment of a best practices and recommendations manual could
increase the chances of integrating the responsible use of RPAS in
conservation and management activities in natural areas. Moreover, some
authors mentioned the lack of commercial operators with sufficient
expertise to carry out such activities (McEvoy, Hall, and McDonald
2016). Also, an optimal trade-off between benefits and environmental
costs should be pursued (Grémillet et al. 2012; Sepúlveda et al. 2010).
By designing quieter, non-polluting and safer components, the impact on
wildlife and ecosystems could be reduced and its objective observation
facilitated. Nonetheless we trust that, as far as further testing be
done, RPAS has great potential to replace more invasive monitoring
techniques, whose reliability is challenged by the potential to induce
conditions of unacceptable stress in wildlife that could ultimately
invalidate the results of the research (Jewell 2013b, Wilson and McMahon
(2006)). This should be consciously taken into account by the many
actors involved in protected areas activities when they are reluctant to
allow RPAS to be an essential part of research and conservation
activities.

\section{Environmental monitoring and decision
support}\label{environmental-monitoring-and-decision-support}

Protected areas are subject to periodic environmental quality control
procedures where RPAS could play a major role. Also, RPAS are suitable
to assist decision making where rapid response is crucial by offering
valuable information at real time to handle natural and man-made
disasters. Wildfires is a major concern in natural parks and is not rare
that RPAS have been put forward to assist in prevention, fighting and
evaluation phases. In most cases, such applications have operational
requirements which eventually are costly. For instance, sophisticated
on-board instruments, gas powered engines for longer endurance and
higher payloads or robotics arms and containers designed to assist
sampling, hold cargo or deliver assistance. RPAS could leverage wildlife
capture procedures by carrying dart guns for chemical immobilization
where otherwise manual approaching free-ranging animals is considered
inefficient or dangerous.

\section{Conclusions}\label{conclusions}

The consolidation of the RPAS as management and research tools in
protected areas is closely linked to the technological development of
the elements associated with the platform and to the establishment of
measures that favorably regulate its use, increasing opportunities in
the sector and stimulating innovation in priority conservation areas.
There are continually improvements in navigation control and flight
autonomy, while we are witnessing the progressive miniaturization and
diversification of sensors along with advances in the field of
artificial intelligence. This rapidly expanding confluence of factors
encourages the emergence of new scenarios with ethical and legal
implications. Most governments have reacted by setting constraints that
could have a negative impact on the capacity to integrate RPAS into the
civilian sphere, despite some progress in this regard. As a result, it
is difficult to foresee the actions that each country will adopt from
now on in an attempt to harmonize the contradictions presented by RPAS,
reason why it is probable that the future of the RPAS in protected areas
is conditioned fundamentally by political and social factors.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-abd-elrahman_development_2005}{}
Abd-Elrahman, Amr, Leonard Pearlstine, and Franklin Percival. 2005.
``Development of Pattern Recognition Algorithm for Automatic Bird .''
\emph{Surv. L. Inf. Sci.} 65 (1): 37.

\hypertarget{ref-Bworld_of_drones2017}{}
America, New. 2017. ``\emph{World of Drones}.''
\url{http://drones.newamerica.org/\#regulations} {[}Accessed: 17 June,
2017{]}.

\hypertarget{ref-van_andel_locating_2015}{}
Andel, Alexander C. van, Serge A. Wich, Christophe Boesch, Lian Pin Koh,
Martha M. Robbins, Joseph Kelly, and Hjalmar S. Kuehl. 2015. ``Locating
Chimpanzee Nests and Identifying Fruiting Trees with an Unmanned Aerial
Vehicle.'' \emph{Am. J. Primatol.} 77 (10): 1122--1134.
doi:\href{https://doi.org/10.1002/ajp.22446}{10.1002/ajp.22446}.

\hypertarget{ref-andrew_semi-automated_2017}{}
Andrew, Margaret E, and Jill M Shephard. 2017. ``Semi-Automated
Detection of Eagle Nests: An Application of Very High-Resolution Image
Data and Advanced Image Analyses to Wildlife Surveys.''
doi:\href{https://doi.org/10.1002/rse2.38}{10.1002/rse2.38}.

\hypertarget{ref-Andrews2014}{}
Andrews, C. 2014. ``\emph{Wildlife Monitoring: Should Uav Drones Be
Banned?}''
\url{https://prod-eandt.theiet.org/content/articles/2014/07/wildlife-monitoring-should-uav-drones-be-banned/}
{[}Accessed: 07 Abril, 2017{]}.

\hypertarget{ref-banzi_sensor_2014}{}
Banzi, Jamali Firmat. 2014. ``A Sensor Based Anti-Poaching System in
Tanzania National Parks.'' \emph{International Journal of Scientific and
Research Publications} 4 (4).

\hypertarget{ref-barasona_unmanned_2014}{}
Barasona, José A., Margarita Mulero-Pázmány, Pelayo Acevedo, Juan J.
Negro, María J. Torres, Christian Gort??zar, and Joaquín Vicente. 2014.
``Unmanned Aircraft Systems for Studying Spatial Abundance of Ungulates:
Relevance to Spatial Epidemiology.'' \emph{PLoS One} 9 (12): 1--17.
doi:\href{https://doi.org/10.1371/journal.pone.0115608}{10.1371/journal.pone.0115608}.

\hypertarget{ref-bareth_low-weight_2015}{}
Bareth, Georg, Helge Aasen, Juliane Bendig, Martin Leon Gnyp, Andreas
Bolten, Andr?s Jung, Ren? Michels, and Jussi Soukkam?ki. 2015.
``Low-Weight and UAV-Based Hyperspectral Full-Frame Cameras for
Monitoring Crops: Spectral Comparison with Portable Spectroradiometer
Measurements.'' \emph{Photogrammetrie - Fernerkundung - Geoinformation}
2015 (1): 69--79.
doi:\href{https://doi.org/10.1127/pfg/2015/0256}{10.1127/pfg/2015/0256}.

\hypertarget{ref-bayram_active_2016}{}
Bayram, Haluk, Krishna Doddapaneni, Nikolaos Stefas, and Volkan Isler.
2016. ``Active Localization of VHF Collared Animals with Aerial
Robots,'' no. 13: 74--75.
doi:\href{https://doi.org/10.1109/COASE.2016.7743503}{10.1109/COASE.2016.7743503}.

\hypertarget{ref-borrelle_will_2017}{}
BORRELLE, SB, and AT FLETCHER. 2017. ``Will Drones Reduce Investigator
Disturbance to Surface-Nesting Seabirds?'' bibtex: borrelle\_will\_2017.
\emph{Marine Ornithology} 45: 89--94.

\hypertarget{ref-bustamante_forest_2015}{}
Bustamante, Luis Antonio Esquivel. 2015. ``Forest Monitoring with Drones
: Application Strategies for Protected Riverine Forest Ecosystems in the
Atlantic Forest of Rio de,'' 96.

\hypertarget{ref-casella_mapping_2017}{}
Casella, Elisa, Antoine Collin, Daniel Harris, Sebastian Ferse, Sonia
Bejarano, Valeriano Parravicini, James L. Hench, and Alessio Rovere.
2017. ``Mapping Coral Reefs Using Consumer-Grade Drones and Structure
from Motion Photogrammetry Techniques.'' \emph{Coral Reefs} 36 (1):
269--275.
doi:\href{https://doi.org/10.1007/s00338-016-1522-0}{10.1007/s00338-016-1522-0}.

\hypertarget{ref-casella_drones_2016}{}
Casella, Elisa, Alessio Rovere, Andrea Pedroncini, Colin P. Stark, Marco
Casella, Marco Ferrari, and Marco Firpo. 2016. ``Drones as Tools for
Monitoring Beach Topography Changes in the Ligurian Sea (NW
Mediterranean).'' \emph{Geo-Marine Letters} 36 (2): 151--163.
doi:\href{https://doi.org/10.1007/s00367-016-0435-9}{10.1007/s00367-016-0435-9}.

\hypertarget{ref-chabot_evaluation_2012}{}
Chabot, Dominique, and David M. Bird. 2012. ``Evaluation of an
Off-the-Shelf Unmanned Aircraft System for Surveying Flocks of Geese.''
\emph{Waterbirds} 35 (1): 170--174.
doi:\href{https://doi.org/10.1675/063.035.0119}{10.1675/063.035.0119}.

\hypertarget{ref-chabot_wildlife_2015}{}
Chabot, Dominique, and David M. Bird. 2015. ``Wildlife Research and
Management Methods in the 21st Century: Where Do Unmanned Aircraft Fit
in?'' \emph{J. Unmanned Veh. Syst.} 3 (4): 137--155.
doi:\href{https://doi.org/10.1139/juvs-2015-0021}{10.1139/juvs-2015-0021}.

\hypertarget{ref-chabot_computer-automated_2016}{}
Chabot, Dominique, and Charles M. Francis. 2016. ``Computer-Automated
Bird Detection and Counts in High-Resolution Aerial Images: A Review.''
\emph{Journal of Field Ornithology} 87 (4): 343--359.
doi:\href{https://doi.org/10.1111/jofo.12171}{10.1111/jofo.12171}.

\hypertarget{ref-christiansen_automated_2014}{}
Christiansen, Peter, Kim A rild Steen, Rasmus N yholm Jørgensen, and
Henrik Karstoft. 2014. ``Automated Detection and Recognition of Wildlife
Using Thermal Cameras.'' \emph{Sensors (Basel).} 14 (8): 13778--13793.
doi:\href{https://doi.org/10.3390/s140813778}{10.3390/s140813778}.

\hypertarget{ref-christie_unmanned_2016}{}
Christie, Katherine S., Sophie L. Gilbert, Casey L. Brown, Michael
Hatfield, and Leanne Hanson. 2016. ``Unmanned Aircraft Systems in
Wildlife Research: Current and Future Applications of a Transformative
Technology.'' \emph{Front. Ecol. Environ.} 14 (5): 241--251.
doi:\href{https://doi.org/10.1002/fee.1281}{10.1002/fee.1281}.

\hypertarget{ref-cliff_online_2015}{}
Cliff, Oliver M, Robert Fitch, Salah Sukkarieh, Debra L Saunders, and
Robert Heinsohn. 2015. ``Online Localization of Radio-Tagged Wildlife
with an Autonomous Aerial Robot System.'' \emph{Robot. Sci. Syst.}, no.
November 2016: 1--9.
doi:\href{https://doi.org/10.15607/RSS.2015.XI.042}{10.15607/RSS.2015.XI.042}.

\hypertarget{ref-colefax_potential_2017}{}
Colefax, Andrew P., Paul A. Butcher, and Brendan P. Kelaher. 2017. ``The
Potential for Unmanned Aerial Vehicles (UAVs) to Conduct Marine Fauna
Surveys in Place of Manned Aircraft.'' \emph{ICES Journal of Marine
Science}, June.
doi:\href{https://doi.org/10.1093/icesjms/fsx100}{10.1093/icesjms/fsx100}.

\hypertarget{ref-Global2017}{}
Consulting, OZYRPAS. 2017. ``\emph{Global Drone Regulations Database}.''
\url{https://www.droneregulations.info/index.html} {[}Accessed: 19 July,
2017{]}.

\hypertarget{ref-cornell_use_2016}{}
Cornell, Dylan, Maryann Herman, and Fernando Ontiveros. 2016. ``Use of a
UAV for Water Sampling to Assist Remote Sensing of Bacterial Flora in
Freshwater Environments.''
\url{http://fisherpub.sjfc.edu/undergraduate_ext_pub/17/}.

\hypertarget{ref-ditmer_bears_2015}{}
Ditmer, Mark A., John B. Vincent, Leland K. Werden, Jessie C. Tanner,
Timothy G. Laske, Paul A. Iaizzo, David L. Garshelis, and John R.
Fieberg. 2015. ``Bears Show a Physiological but Limited Behavioral
Response to Unmanned Aerial Vehicles.'' \emph{Curr. Biol.} 25 (17):
2278--2283.
doi:\href{https://doi.org/10.1016/j.cub.2015.07.024}{10.1016/j.cub.2015.07.024}.

\hypertarget{ref-dudley_guidelines_2008}{}
Dudley, Nigel. 2008. \emph{Guidelines for Applying Protected Area
Management Categories}. bibtex: dudley\_guidelines\_2008. IUCN.

\hypertarget{ref-duffy_waging_2014}{}
Duffy, Rosaleen. 2014. ``Waging a War to Save Biodiversity: The Rise of
Militarized Conservation.'' \emph{Int. Aff.} 90 (4): 819--834.
doi:\href{https://doi.org/10.1111/1468-2346.12142}{10.1111/1468-2346.12142}.

\hypertarget{ref-dulava_environmental_2015}{}
Dulava, Sharon, William T. Bean, and Orien M. W. Richmond. 2015.
``ENVIRONMENTAL REVIEWS AND CASE STUDIES: Applications of Unmanned
Aircraft Systems (UAS) for Waterbird Surveys.'' \emph{Environ. Pract.}
17 (3): 201--210.
doi:\href{https://doi.org/10.1017/S1466046615000186}{10.1017/S1466046615000186}.

\hypertarget{ref-durban_photogrammetry_2015}{}
Durban, J W, H Fearnbach, W L Perryman, and D J Leroi. 2015.
``Photogrammetry of Killer Whales Using a Small Hexacopter Launched at
Sea.'' \emph{J. Unmanned Veh. Syst.} 3 (June): 1--5.
doi:\href{https://doi.org/dx.doi.org/10.1139/juvs-2015-0020}{dx.doi.org/10.1139/juvs-2015-0020}.

\hypertarget{ref-fornace_mapping_2014}{}
Fornace, Kimberly M., Chris J. Drakeley, Timothy William, Fe Espino, and
Jonathan Cox. 2014. ``Mapping Infectious Disease Landscapes: Unmanned
Aerial Vehicles and Epidemiology.'' \emph{Trends Parasitol.}, October,
1--6.
doi:\href{https://doi.org/10.1016/j.pt.2014.09.001}{10.1016/j.pt.2014.09.001}.

\hypertarget{ref-franco_five_2016}{}
Franco, Antonio Di, Pierre Thiriet, Giuseppe Di Carlo, Charalampos
Dimitriadis, Patrice Francour, Nicolas L Gutiérrez, Alain Jeudy De
Grissac, et al. 2016. ``Five Key Attributes Can Increase Marine
Protected Areas Performance for Small-Scale Fisheries Management.''
\emph{Nat. Publ. Gr.}, no. November: 1--9.
doi:\href{https://doi.org/10.1038/srep38135}{10.1038/srep38135}.

\hypertarget{ref-van_gemert_nature_2015}{}
Gemert, Jan C. van, Camiel R. Verschoor, Pascal Mettes, Kitso Epema,
Lian Pin Koh, and Serge Wich. 2015. ``Nature Conservation Drones for
Automatic Localization and Counting of Animals.'' \emph{Lect. Notes
Comput. Sci. (Including Subser. Lect. Notes Artif. Intell. Lect. Notes
Bioinformatics)} 8925: 255--270.
doi:\href{https://doi.org/10.1007/978-3-319-16178-5_17}{10.1007/978-3-319-16178-5\_17}.

\hypertarget{ref-gini_aerial_2012}{}
Gini, R., D. Passoni, L. Pinto, and G. Sona. 2012. ``Aerial Images from
an Uav System: 3D Modeling and Tree Species Classification in a Park
Area.'' \emph{ISPRS - Int. Arch. Photogramm. Remote Sens. Spat. Inf.
Sci.} XXXIX-B1 (September): 361--366.
doi:\href{https://doi.org/10.5194/isprsarchives-XXXIX-B1-361-2012}{10.5194/isprsarchives-XXXIX-B1-361-2012}.

\hypertarget{ref-goebel_small_2015}{}
Goebel, Michael E., Wayne L. Perryman, Jefferson T. Hinke, Douglas J.
Krause, Nancy A. Hann, Steve Gardner, and Donald J. LeRoi. 2015. ``A
Small Unmanned Aerial System for Estimating Abundance and Size of
Antarctic Predators.'' \emph{Polar Biol.} 38 (5): 619--630.
doi:\href{https://doi.org/10.1007/s00300-014-1625-4}{10.1007/s00300-014-1625-4}.

\hypertarget{ref-gonzalez_unmanned_2016}{}
Gonzalez, Luis F., Glen A. Montes, Eduard Puig, Sandra Johnson, Kerrie
Mengersen, and Kevin J. Gaston. 2016. ``Unmanned Aerial Vehicles (UAVs)
and Artificial Intelligence Revolutionizing Wildlife Monitoring and
Conservation.'' \emph{Sensors (Switzerland)} 16 (1).
doi:\href{https://doi.org/10.3390/s16010097}{10.3390/s16010097}.

\hypertarget{ref-gremillet_robots_2012}{}
Grémillet, David, William Puech, Véronique Garçon, Thierry Boulinier,
and Yvon Le Maho. 2012. ``Robots in Ecology: Welcome to the Machine.''
\emph{Open J. Ecol.} 02 (2): 49--57.
doi:\href{https://doi.org/10.4236/oje.2012.22006}{10.4236/oje.2012.22006}.

\hypertarget{ref-hansen_applying_2016}{}
Hansen, Andreas Skriver. 2016. ``Applying Visitor Monitoring Methods in
Coastal and Marine Areas -- Some Learnings and Critical Reflections from
Sweden.'' \emph{Scand. J. Hosp. Tour.} 2250 (June): 1--18.
doi:\href{https://doi.org/10.1080/15022250.2016.1155481}{10.1080/15022250.2016.1155481}.

\hypertarget{ref-hodgson_unmanned_2013}{}
Hodgson, Amanda, Natalie Kelly, and David Peel. 2013. ``Unmanned Aerial
Vehicles (UAVs) for Surveying Marine Fauna: A Dugong Case Study.''
\emph{PLoS One} 8 (11): 1--15.
doi:\href{https://doi.org/10.1371/journal.pone.0079556}{10.1371/journal.pone.0079556}.

\hypertarget{ref-hodgson_unmanned_2017}{}
Hodgson, Amanda, David Peel, and Natalie Kelly. 2017. ``Unmanned Aerial
Vehicles for Surveying Marine Fauna: Assessing Detection Probability.''
\emph{Ecological Applications}.
\url{http://onlinelibrary.wiley.com/doi/10.1002/eap.1519/full}.

\hypertarget{ref-hodgson_precision_2016}{}
Hodgson, Jarrod C, Shane M Baylis, Rowan Mott, Ashley Herrod, and Rohan
H Clarke. 2016. ``Precision Wildlife Monitoring Using Unmanned Aerial
Vehicles.'' \emph{Sci. Rep.} 6 (March): 22574.
doi:\href{https://doi.org/10.1038/srep22574}{10.1038/srep22574}.

\hypertarget{ref-hodgson_best_2016}{}
Hodgson, Jarrod C., and Lian Pin Koh. 2016. ``Best Practice for
Minimising Unmanned Aerial Vehicle Disturbance to Wildlife in Biological
Field Research.'' \emph{Curr. Biol.} 26 (10).
doi:\href{https://doi.org/10.1016/j.cub.2016.04.001}{10.1016/j.cub.2016.04.001}.

\hypertarget{ref-ivosevic_use_2015}{}
Ivošević, Bojana, Yong Gu Han, Youngho Cho, and Ohseok Kwon. 2015. ``The
Use of Conservation Drones in Ecology and Wildlife Research.'' \emph{J.
Ecol. Environ.} 38 (1): 113--118.
doi:\href{https://doi.org/10.5141/ecoenv.2015.012}{10.5141/ecoenv.2015.012}.

\hypertarget{ref-jain_unmanned_2013}{}
Jain, Mukesh. 2013. ``Unmanned Aerial Survey of Elephants.'' \emph{PLoS
One}.
doi:\href{https://doi.org/10.1371/\%20journal.pone.0054700}{10.1371/ journal.pone.0054700}.

\hypertarget{ref-jewell_effect_2013-1}{}
Jewell, Zoe. 2013a. ``Effect of Monitoring Technique on Quality of
Conservation Science: Ethics and Science in Conservation.''
\emph{Conservation Biology} 27 (3): 501--508.
doi:\href{https://doi.org/10.1111/cobi.12066}{10.1111/cobi.12066}.

\hypertarget{ref-jewell_effect_2013}{}
Jewell, Zoe. 2013b. ``Effect of Monitoring Technique on Quality of
Conservation Science: Ethics and Science in Conservation.''
\emph{Conservation Biology} 27 (3): 501--508.
doi:\href{https://doi.org/10.1111/cobi.12066}{10.1111/cobi.12066}.

\hypertarget{ref-king_will_2014}{}
King, Lisa M. 2014. ``Will Drones Revolutionise Ecotourism?'' \emph{J.
Ecotourism} 13 (1): 85--92.
doi:\href{https://doi.org/10.1080/14724049.2014.948448}{10.1080/14724049.2014.948448}.

\hypertarget{ref-koh_dawn_2012}{}
Koh, Lian Pin, and Serge A. Wich. 2012. ``Dawn of Drone Ecology:
Low-Cost Autonomous Aerial Vehicles for Conservation.'' \emph{Trop.
Conserv. Sci.} 5 (2): 121--132.
doi:\href{https://doi.org/WOS:000310846600002}{WOS:000310846600002}.

\hypertarget{ref-koski_evaluation_2009}{}
Koski, William R., Travis Allen, Darren Ireland, Greg Buck, Paul R.
Smith, A. Michael Macrander, Melissa A. Halick, Chris Rushing, David J.
Sliwa, and Trent L. McDonald. 2009. ``Evaluation of an Unmanned Airborne
System for Monitoring Marine Mammals.'' \emph{Aquatic Mammals} 35 (3):
347--357.
doi:\href{https://doi.org/10.1578/AM.35.3.2009.347}{10.1578/AM.35.3.2009.347}.

\hypertarget{ref-koski_evaluation_2015}{}
Koski, William R., Gayan Gamage, Andrew R. Davis, Tony Mathews, Bernard
LeBlanc, and Steven H. Ferguson. 2015. ``Evaluation of UAS for
Photographic Re-Identification of Bowhead Whales, Balaena Mysticetus.''
\emph{J. Unmanned Veh. Syst.} 3 (1): 22--29.
doi:\href{https://doi.org/10.1139/juvs-2014-0014}{10.1139/juvs-2014-0014}.

\hypertarget{ref-korner_autonomous_2010}{}
Körner, Fabian, Raphael Speck, Ali Haydar, and Salah Sukkarieh. 2010.
``Autonomous Airborne Wildlife Tracking Using Radio Signal Strength,''
107--112.

\hypertarget{ref-lancia_estimating_2005}{}
Lancia, Richard A, William L Kendall, Kenneth H Pollock, and James D
Nichols. 2005. ``Estimating the Number of Animals in Wildlife
Populations.'' bibtex: lancia\_estimating\_2005.

\hypertarget{ref-leary_drones_2017}{}
Leary, David. 2017. ``Drones on Ice: An Assessment of the Legal
Implications of the Use of Unmanned Aerial Vehicles in Scientific
Research and by the Tourist Industry in Antarctica.'' \emph{Polar
Record}, May, 1--15.
doi:\href{https://doi.org/10.1017/S0032247417000262}{10.1017/S0032247417000262}.

\hypertarget{ref-leonardo_miniature_2013}{}
Leonardo, Miguel, Austin Jensen, Calvin Coopmans, Mac McKee, and
YangQuan Chen. 2013. ``A Miniature Wildlife Tracking UAV Payload System
Using Acoustic Biotelemetry.'' \emph{Proc. ASME Int. Des. Eng. Tech.
Conf. Comput. Inf. Eng. Conf.}, no. July 2015.
doi:\href{https://doi.org/10.1115/DETC2013-13267}{10.1115/DETC2013-13267}.

\hypertarget{ref-lhoest_how_2015}{}
Lhoest, S., J. Linchant, S. Quevauvillers, C. Vermeulen, and P. Lejeune.
2015. ``How Many Hippos (Homhip): Algorithm for Automatic Counts of
Animals with Infra-Red Thermal Imagery from UAV.'' \emph{Int. Arch.
Photogramm. Remote Sens. Spat. Inf. Sci. - ISPRS Arch.} 40 (3):
355--362.
doi:\href{https://doi.org/10.5194/isprsarchives-XL-3-W3-355-2015}{10.5194/isprsarchives-XL-3-W3-355-2015}.

\hypertarget{ref-linchant_are_2015}{}
Linchant, Julie, Jonathan Lisein, Jean Semeki, Philippe Lejeune, and
Cédric Vermeulen. 2015. ``Are Unmanned Aircraft Systems (UASs) the
Future of Wildlife Monitoring? A Review of Accomplishments and
Challenges.'' \emph{Mamm. Rev.} 45 (4): 239--252.
doi:\href{https://doi.org/10.1111/mam.12046}{10.1111/mam.12046}.

\hypertarget{ref-lisein_discrimination_2015}{}
Lisein, Jonathan, Adrien Michez, Hugues Claessens, and Philippe Lejeune.
2015. ``Discrimination of Deciduous Tree Species from Time Series of
Unmanned Aerial System Imagery.'' \emph{PLoS One} 10 (11).
doi:\href{https://doi.org/10.1371/journal.pone.0141006}{10.1371/journal.pone.0141006}.

\hypertarget{ref-lobermeier_mitigating_2015}{}
Lobermeier, Scott, Matthew Moldenhauer, Christopher Peter, Luke
Slominski, Richard Tedesco, Marcus Meer, James Dwyer, Richard Harness,
and Andrew Stewart. 2015. ``Mitigating Avian Collision with Power Lines:
A Proof of Concept for Installation of Line Markers via Unmanned Aerial
Vehicle.'' \emph{J. Unmanned Veh. Syst.} 3 (4): 252--258.
doi:\href{https://doi.org/10.1139/juvs-2015-0009}{10.1139/juvs-2015-0009}.

\hypertarget{ref-longmore_adapting_2017}{}
Longmore, S. N., R. P. Collins, S. Pfeifer, S. E. Fox, M.
Mulero-Pazmany, F. Bezombes, A. Goodwind, M. de Juan Ovelar, J. H.
Knapen, and S. A. Wich. 2017. ``Adapting Astronomical Source Detection
Software to Help Detect Animals in Thermal Images Obtained by Unmanned
Aerial Systems'' 00 (0): 1--16.
doi:\href{https://doi.org/10.1080/01431161.2017.1280639}{10.1080/01431161.2017.1280639}.

\hypertarget{ref-margarita_mulero-pazmany_juan_jose_negro_low_2014}{}
Margarita Mulero-Pázmány, Miguel Ferrer, Juan José Negro. 2014. ``A Low
Cost Way for Assessing Bird Risk Hazards in Power Lines: Fixed-Wing
Small Unmanned Aircraft Systems'' 2.

\hypertarget{ref-martin_estimating_2012}{}
Martin, Julien, Holly H. Edwards, Matthew A. Burgess, H. Franklin
Percival, Daniel E. Fagan, Beth E. Gardner, Joel G. Ortega-Ortiz, Peter
G. Ifju, Brandon S. Evers, and Thomas J. Rambo. 2012. ``Estimating
Distribution of Hidden Objects with Drones: From Tennis Balls to
Manatees.'' \emph{PLoS One} 7 (6): 1--8.
doi:\href{https://doi.org/10.1371/journal.pone.0038882}{10.1371/journal.pone.0038882}.

\hypertarget{ref-maza_experimental_2011}{}
Maza, Iv?n, Fernando Caballero, Jes?s Capit?n, J. R. Mart?nez-de-Dios,
and An?bal Ollero. 2011. ``Experimental Results in Multi-UAV
Coordination for Disaster Management and Civil Security Applications.''
\emph{Journal of Intelligent \& Robotic Systems} 61 (1): 563--585.
doi:\href{https://doi.org/10.1007/s10846-010-9497-5}{10.1007/s10846-010-9497-5}.

\hypertarget{ref-mccaldin_use_2015}{}
McCaldin, Guy, Michael Johnston, and Andrew Rieker. 2015. \emph{Use of
Unmanned Aircraft Systems to Assist with Decision Support for Land
Managers on Christmas Island (Indian Ocean)}. October. Australia: V-TOL
Aerospace; Department of parks; Wildlife, Western Australia.

\hypertarget{ref-mcevoy_evaluation_2016}{}
McEvoy, John F., Graham P. Hall, and Paul G. McDonald. 2016.
``Evaluation of Unmanned Aerial Vehicle Shape, Flight Path and Camera
Type for Waterfowl Surveys: Disturbance Effects and Species
Recognition.'' \emph{PeerJ} 4 (March): e1831.
doi:\href{https://doi.org/10.7717/peerj.1831}{10.7717/peerj.1831}.

\hypertarget{ref-miyamoto_use_2004}{}
Miyamoto, Michiru, Kunihiko Yoshino, Toshihide Nagano, Tomoyasu Ishida,
and Yohei Sato. 2004. ``Use of Balloon Aerial Photography for
Classification of Kushiro Wetland Vegetation, Northeastern Japan.''
\emph{Wetlands} 24 (3): 701--710.
doi:\href{https://doi.org/10.1672/0277-5212(2004)024\%5B0701:UOBAPF\%5D2.0.CO;2}{10.1672/0277-5212(2004)024{[}0701:UOBAPF{]}2.0.CO;2}.

\hypertarget{ref-mulero-pazmany_margarita_aeromab_2011}{}
Mulero-Pázmány Margarita, Negro JJ. 2011. ``AEROMAB Small UAS for
Montagu's Harrier's Circus Pygargus Nests Monitoring.'' \emph{AEROMAB
Small UAS Montagu's Harrier's Circus Pygargus Nests Monit. RED UAS
Intenational Congr. Univ. Eng. Seville, Spain. December 2011.}

\hypertarget{ref-mulero-pazmany_unmanned_2015}{}
Mulero-Pázmány, Margarita, Jose Ángel Barasona, Pelayo Acevedo, Joaquín
Vicente, and Juan José Negro. 2015. ``Unmanned Aircraft Systems
Complement Biologging in Spatial Ecology Studies.'' \emph{Ecol. Evol.} 5
(21): 4808--4818.
doi:\href{https://doi.org/10.1002/ece3.1744}{10.1002/ece3.1744}.

\hypertarget{ref-mulero-pazmany_unmanned_2017}{}
Mulero-Pázmány, Margarita, Susanne Jenni-Eiermann, Nicolas Strebel,
Thomas Sattler, Juan José Negro, and Zulima Tablado. 2017. ``Unmanned
Aircraft Systems as a New Source of Disturbance for Wildlife: A
Systematic Review.'' \emph{PloS One} 12 (6): e0178448.
\url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0178448}.

\hypertarget{ref-mulero-pazmany_remotely_2014}{}
Mulero-Pázmány, Margarita, Roel Stolper, L. D. Van Essen, Juan J. Negro,
and Tyrell Sassen. 2014. ``Remotely Piloted Aircraft Systems as a
Rhinoceros Anti-Poaching Tool in Africa.'' \emph{PLoS One} 9 (1): 1--10.
doi:\href{https://doi.org/10.1371/journal.pone.0083873}{10.1371/journal.pone.0083873}.

\hypertarget{ref-mullerova_does_2016}{}
Müllerová, Jana, Josef Brůna, Dvořák, Peter, Tomáš Bartaloš, and
Michaela Vítková. 2016. ``DOES THE DATA RESOLUTION/ORIGIN MATTER?
SATELLITE, AIRBORNE AND UAV IMAGERY TO TACKLE PLANT INVASIONS.''
\emph{ISPRS - International Archives of the Photogrammetry, Remote
Sensing and Spatial Information Sciences} XLI-B7 (June): 903--908.
doi:\href{https://doi.org/10.5194/isprsarchives-XLI-B7-903-2016}{10.5194/isprsarchives-XLI-B7-903-2016}.

\hypertarget{ref-nugraha_urgency_2016}{}
Nugraha, Ridha Aditya, Deepika Jeyakodi, and Thitipon Mahem. 2016.
``Urgency for Legal Framework on Drones : Lessons for Indonesia , India
, and Thailand.'' \emph{Indones. Law Rev.} 6 (2): 137--157.

\hypertarget{ref-olivares-mendez_noah_2014}{}
Olivares-Mendez, Miguel A., Tegawendé F. Bissyandé, Kannan Somasundar,
Jacques Klein, Holger Voos, and Yves Le Traon. 2014. ``The NOAH Project:
Giving a Chance to Threatened Species in Africa with UAVs.'' \emph{Lect.
Notes Inst. Comput. Sci. Soc. Telecommun. Eng. LNICST} 135 LNICST:
198--208.
doi:\href{https://doi.org/10.1007/978-3-319-08368-1_24}{10.1007/978-3-319-08368-1\_24}.

\hypertarget{ref-perroy_assessing_2017}{}
Perroy, Ryan L., Timo Sullivan, and Nathan Stephenson. 2017. ``Assessing
the Impacts of Canopy Openness and Flight Parameters on Detecting a
Sub-Canopy Tropical Invasive Plant Using a Small Unmanned Aerial
System.'' \emph{ISPRS J. Photogramm. Remote Sens.} 125: 174--183.
doi:\href{https://doi.org/10.1016/j.isprsjprs.2017.01.018}{10.1016/j.isprsjprs.2017.01.018}.

\hypertarget{ref-pomeroy_assessing_2015}{}
Pomeroy, P., L. O'Connor, and P. Davies. 2015. ``Assessing Use of and
Reaction to Unmanned Aerial Systems in Gray and Harbor Seals During
Breeding and Molt in the UK \(^{\textrm{1}}\).'' \emph{Journal of
Unmanned Vehicle Systems} 3 (3): 102--113.
doi:\href{https://doi.org/10.1139/juvs-2015-0013}{10.1139/juvs-2015-0013}.

\hypertarget{ref-puttock_aerial_2015}{}
Puttock, A.K., A.M. Cunliffe, K. Anderson, and R.E. Brazier. 2015.
``Aerial Photography Collected with a Multirotor Drone Reveals Impact of
Eurasian Beaver Reintroduction on Ecosystem Structure 1.'' \emph{J.
Unmanned Veh. Syst.} 3 (3): 123--130.
doi:\href{https://doi.org/10.1139/juvs-2015-0005}{10.1139/juvs-2015-0005}.

\hypertarget{ref-rodriguez_eye_2012}{}
Rodríguez, Airam, Juan J. Negro, Mara Mulero, Carlos Rodríguez, Jesús
Hernández-Pliego, and Javier Bustamante. 2012. ``The Eye in the Sky:
Combined Use of Unmanned Aerial Systems and GPS Data Loggers for
Ecological Research and Conservation of Small Birds.'' \emph{PLoS One} 7
(12).
doi:\href{https://doi.org/10.1371/journal.pone.0050336}{10.1371/journal.pone.0050336}.

\hypertarget{ref-sabella_preliminary_2017}{}
Sabella, Giorgio, Fabio Massimo Viglianisi, Sergio Rotondi, and
Filadelfo Brogna. 2017. ``Preliminary Observations on the Use of Drones
in the Environmental Monitoring and in the Management of Protected
Areas. the Case Study of `RNO Vendicari', Syracuse (Italy).'' bibtex:
sabella\_preliminary\_2017. \emph{Biodiversity Journal} 8: 79--86.
\url{http://www.biodiversityjournal.com/pdf/8(1)_79-86.pdf}.

\hypertarget{ref-sarda-palomera_fine-scale_2012}{}
Sardà-Palomera, Francesc, Gerard Bota, Carlos Viñolo, Oriol Pallarés,
Víctor Sazatornil, Lluís Brotons, Spartacus Gomáriz, and Francesc Sardà.
2012. ``Fine-Scale Bird Monitoring from Light Unmanned Aircraft
Systems.'' \emph{Ibis (Lond. 1859).} 154 (1): 177--183.
doi:\href{https://doi.org/10.1111/j.1474-919X.2011.01177.x}{10.1111/j.1474-919X.2011.01177.x}.

\hypertarget{ref-schmale_development_2008}{}
Schmale, DG, Benjamin R. Dingus, and Charles Reinholtz. 2008.
``Development and Application of an Autonomous Unmanned Aerial Vehicle
for Precise Aerobiological Sampling Above Agricultural Fields.''
\emph{J. F. Robot.} 25 (3): 133--147.
doi:\href{https://doi.org/10.1002/rob}{10.1002/rob}.

\hypertarget{ref-schwarzbach_remote_2014}{}
Schwarzbach, Marc, Maximilian Laiacker, Margarita Mulero-Pazmany, and
Konstantin Kondak. 2014. ``Remote Water Sampling Using Flying Robots.''
\emph{2014 Int. Conf. Unmanned Aircr. Syst. ICUAS 2014 - Conf. Proc.},
72--76.
doi:\href{https://doi.org/10.1109/ICUAS.2014.6842240}{10.1109/ICUAS.2014.6842240}.

\hypertarget{ref-scobie_wildlife_2016}{}
Scobie, Corey A., and Chris H. Hugenholtz. 2016. ``Wildlife Monitoring
with Unmanned Aerial Vehicles: Quantifying Distance to Auditory
Detection: UAV Sound and Wildlife Aural Detection.'' \emph{Wildlife
Society Bulletin} 40 (4): 781--785.
doi:\href{https://doi.org/10.1002/wsb.700}{10.1002/wsb.700}.

\hypertarget{ref-sepulveda_review_2010}{}
Sepúlveda, Alejandra, Mathias Schluep, Fabrice G. Renaud, Martin
Streicher, Ruediger Kuehr, Christian Hagelüken, and Andreas C. Gerecke.
2010. ``A Review of the Environmental Fate and Effects of Hazardous
Substances Released from Electrical and Electronic Equipments During
Recycling: Examples from China and India.'' \emph{Environmental Impact
Assessment Review} 30 (1): 28--41.
doi:\href{https://doi.org/10.1016/j.eiar.2009.04.001}{10.1016/j.eiar.2009.04.001}.

\hypertarget{ref-soriano_rf-based_2009}{}
Soriano, P, F Caballero, and A Ollero. 2009. ``RF-Based Particle Filter
Localization for Wildlife Tracking by Using an UAV.'' \emph{40 Th Int.
Symp. Robot.}, 239--244.
\url{http://grvc.us.es/publica/congresosint/documentos/isr_soriano.pdf}.

\hypertarget{ref-stocker_review_2017}{}
Stöcker, Claudia, Rohan Bennett, Francesco Nex, Markus Gerke, and Jaap
Zevenbergen. 2017. ``Review of the Current State of UAV Regulations.''
\emph{Remote Sensing} 9 (5): 459.
doi:\href{https://doi.org/10.3390/rs9050459}{10.3390/rs9050459}.

\hypertarget{ref-szantoi_mapping_2017}{}
Szantoi, Zoltan, Scot E. Smith, Giovanni Strona, Lian Pin Koh, and Serge
A. Wich. 2017. ``Mapping Orangutan Habitat and Agricultural Areas Using
Landsat OLI Imagery Augmented with Unmanned Aircraft System Aerial
Photography.'' \emph{Int. J. Remote Sens.} 38 (8): 1--15.
doi:\href{https://doi.org/10.1080/01431161.2017.1280638}{10.1080/01431161.2017.1280638}.

\hypertarget{ref-AUVSI2017}{}
UAVSI. 2017. ``\emph{Association for Unmanned Vehicle Systems
International}.'' \url{http://www.auvsi.org} {[}Accessed: 27 July,
2017{]}.

\hypertarget{ref-USGS2014}{}
USGS. 2014. ``\emph{US Geological Survey National Unmanned Aircraft
Systems Project}.''
\href{http://\%20rmgsc.cr.usgs.gov/UAS}{http:// rmgsc.cr.usgs.gov/UAS}
{[}Accessed: 13 April, 2017{]}.

\hypertarget{ref-van_tilburg_first_2017}{}
Van Tilburg, Christopher. 2017. ``First Report of Using Portable
Unmanned Aircraft Systems (Drones) for Search and Rescue.''
\emph{Wilderness \& Environmental Medicine}.
\url{http://www.sciencedirect.com/science/article/pii/S1080603217300042}.

\hypertarget{ref-vas_approaching_2015}{}
Vas, E., A. Lescroel, O. Duriez, G. Boguszewski, and D. Gremillet. 2015.
``Approaching Birds with Drones: First Experiments and Ethical
Guidelines.'' \emph{Biol. Lett.} 11 (2): 20140754--20140754.
doi:\href{https://doi.org/10.1098/rsbl.2014.0754}{10.1098/rsbl.2014.0754}.

\hypertarget{ref-ventura_low-cost_2016}{}
Ventura, Daniele, Michele Bruno, Giovanna Jona Lasinio, Andrea
Belluscio, and Giandomenico Ardizzone. 2016. ``A Low-Cost Drone Based
Application for Identifying and Mapping of Coastal Fish Nursery
Grounds.'' \emph{Estuar. Coast. Shelf Sci.} 171.
doi:\href{https://doi.org/10.1016/j.ecss.2016.01.030}{10.1016/j.ecss.2016.01.030}.

\hypertarget{ref-weissensteiner_low-budget_2015}{}
Weissensteiner, M. H., J. W. Poelstra, and J. B. W. Wolf. 2015.
``Low-Budget Ready-to-Fly Unmanned Aerial Vehicles: An Effective Tool
for Evaluating the Nesting Status of Canopy-Breeding Bird Species.''
bibtex: weissensteiner\_low-budget\_2015. \emph{Journal of Avian
Biology} 46 (4): 425--430.
doi:\href{https://doi.org/10.1111/jav.00619}{10.1111/jav.00619}.

\hypertarget{ref-wich_preliminary_2016}{}
Wich, Serge, David Dellatore, Max Houghton, Rio Ardi, and Lian Pin Koh.
2016. ``A Preliminary Assessment of Using Conservation Drones for
Sumatran Orang-Utan (Pongo Abelii) Distribution and Density.'' \emph{J.
Unmanned Veh. Syst.} 4 (1): 45--52.
doi:\href{https://doi.org/10.1139/juvs-2015-0015}{10.1139/juvs-2015-0015}.

\hypertarget{ref-wilson_feasibility_2017}{}
Wilson, Andrew M, Janine Barr, and Megan Zagorski. 2017. ``The
Feasibility of Counting Songbirds Using Unmanned Aerial Vehicles.''
\emph{Auk} 134 (2): 350--362.
doi:\href{https://doi.org/10.1642/AUK-16-216.1}{10.1642/AUK-16-216.1}.

\hypertarget{ref-wilson_measuring_2006}{}
Wilson, Rory P., and Clive R. McMahon. 2006. ``Measuring Devices on Wild
Animals: What Constitutes Acceptable Practice?'' \emph{Frontiers in
Ecology and the Environment} 4 (3): 147--154.
doi:\href{https://doi.org/10.1890/1540-9295(2006)004\%5B0147:MDOWAW\%5D2.0.CO;2}{10.1890/1540-9295(2006)004{[}0147:MDOWAW{]}2.0.CO;2}.

\hypertarget{ref-wulder_high_2004}{}
Wulder, Michael A, Ronald J Hall, Nicholas C Coops, and Steven E
Franklin. 2004. ``High Spatial Resolution Remotely Sensed Data for
Ecosystem Characterization'' 54 (6): 511--521.
doi:\href{https://doi.org/10.1641/0006-3568(2004)054}{10.1641/0006-3568(2004)054}.

\hypertarget{ref-zahawi_using_2015}{}
Zahawi, Rakan A., Jonathan P. Dandois, Karen D. Holl, Dana Nadwodny, J.
Leighton Reid, and Erle C. Ellis. 2015. ``Using Lightweight Unmanned
Aerial Vehicles to Monitor Tropical Forest Recovery.'' \emph{Biol.
Conserv.} 186 (June): 287--295.
doi:\href{https://doi.org/10.1016/j.biocon.2015.03.031}{10.1016/j.biocon.2015.03.031}.

\hypertarget{ref-zaman_use_2011}{}
Zaman, Bushra, Austin M. Jensen, and Mac McKee. 2011. ``Use of
High-Resolution Multispectral Imagery Acquired with an Autonomous
Unmanned Aerial Vehicle to Quantify the Spread of an Invasive Wetlands
Species.'' \emph{Int. Geosci. Remote Sens. Symp.}, 803--806.
doi:\href{https://doi.org/10.1109/IGARSS.2011.6049252}{10.1109/IGARSS.2011.6049252}.

\hypertarget{ref-zang_investigating_2012}{}
Zang, Wenqian, Jiayuan Lin, Yangchun Wang, and Heping Tao. 2012.
``Investigating Small-Scale Water Pollution with UAV Remote Sensing
Technology.'' In \emph{World Automation Congress (WAC), 2012}, 1--4.
IEEE. \url{http://ieeexplore.ieee.org/abstract/document/6321515/}.

\newpage
\singlespacing 
\end{document}
